### 2008-04-15 viewer_paper/paper.tex
Paper for ImageJ 3D Viewer and target in the Makefile. (ImageJ_3D_Viewer.pdf)

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{subfig}

%opening
\title{A 3D viewer for ImageJ}
\author{B. Schmid, J. Schindelin, A. Cardona and M. Heisenberg}

\begin{document}

\maketitle
\setlength{\parskip}{0.3cm}

\begin{abstract}
In this paper, we introduce a tool for visualizing 3D datasets. Microscopy techniques in biology and medicine, like confocal microscopy, EM, MRI etc, often provide a stack of 2D images. For the investigation of these images, however, the two-dimensional view of the single slices is often not sufficient. We therefore developped a software for 3D visualisation, which allows to display the stacks optionally as surfaces, volume renderings or orthoslices. The software focusses on interactive, hardware-accelerated display of the stacks. Additionally, it offers possibilities for animation and movie recording, as well as an elaborate way to handle transformations of displayed bodies.
Three major aspects make this tool outstanding against existing software in the targeted field of usage:
\begin{itemize}
\item It is freely available and extensible.
\item It was developped for simplicity of usage.
\item It is incorporated into ImageJ \cite{imagej}, a prominent image processing software for microscopy data.
\end{itemize}
\end{abstract}

\section{Introduction}
Medical and biological imaging techniques often acquire 3D datasets in the form of stacks of 2D images. Confocal microscopy for example can be used to cut a \textit{Drosophila} brain virtually into slices of a few $\mu$m thickness and scans it. For many areas of research and application, the 2D view is not sufficient, however. For anatomists, e.g., it is vital to get an exact representation of the form and spatial relations of the tissue under study.

While there exists software for 3D visualisation, it suffers from several drawbacks: 3D software providing a comprehensive set of features is often proprietary and expensive, complicated in its usage or not suited for biological data. The tool we present does not share these shortcomings. While not restricted to this field of application, it was primarily designed for microscopic data. Its most relevant advantage however is its incorporation into ImageJ, which is one of the most wide-spread image processing packages in the field of microscopy. Users of ImageJ who already benefit from its numerous possibilities and plugins to process microscopic images, like filtering and deconvolution \cite{mbf}, can now visualize their data in 3D seemlessly, without the need of another software.

This paper is organized as follows: We first introduce briefly the technologies used for the implementation of the software. Subsequently, we show how the surface, volume and orthoslice views are created, as well as more information about the features of this software. Finally, more details about the implementation is provided.

\section{Technologies}
The technologies which were used in this work are ImageJ, Java and Java 3D.

\textbf{ImageJ} \cite{imagej} is used as underlying platform. There are several reasons for choosing ImageJ for this task: Using an image processing platform in general has the advantage that the developper does not need to take care about low-level actions like opening/saving images etc, since functionality like that is given by the chosen platform. The choice fell on ImageJ since it is by far the most widely used image processing software in the biological and medical field \cite{mbf}. Several reasons account for this:
\begin{itemize}
\item It is freely available and open source.
\item It consists of a compact core covering basic functionality, and is easily extendible to the user's needs by plugins.
\item There exists a active and competent community, which both serves as support via online forums and as developers for plugins of which hundreds can be found on the internet.
\end{itemize}
Using ImageJ facilitated our development as well as it facilitates the usage of the 3D viewer by scientists who are already familiar with ImageJ.

\textbf{Java} \cite{java} is a programming language which is platform independent. This means that our software runs on windows, unix and mac machines. ImageJ is written in Java.

\textbf{Java 3D.} To guarantee interactive 3D visualisation, hardware capabilities must be used optimally. Many 3D rendering tasks can nowadays be fullfilled by graphics cards, which frees the CPU and accelerates operations drastically. Java 3D \cite{java3d} is a technology which allows to use these hardware capabilities from within Java. An alternative to Java 3D would have been JOGL, a Java binding for OpenGL. However, in our eyes Java 3D is the more promising technology.

\section{3D Visualisation}
3D datasets can be represented in different ways. The options we implemented in this work are surfaces, volume renderings and orthoslices.
Starting point is a set of 2D slices which contain volumetric data, e.g. acquired by confocal microscopy. A stack like this can optionally be displayed as a three dimensional surface, orthoslices or as a volume rendering. These possibilities are described in more detail in the following paragraphs.
\subsection{Surfaces}
A surface is generally understood as the border between a volumetric body and its environment. Given a volumetric 3D dataset, constructing a corresponding surface is an ill-defined problem. In computer graphics, one often uses isosurfaces for this purpose. Isosurfaces assume that there is an intensity threshold which separates a body from its background. This border defines the isosurface, it therefore has the same intensity everywhere, which is also called its isovalue.

The most used representation of surfaces in computer graphics are polygon meshes, in particular triangular meshes. Java 3D provides methods to display triangle meshes and is capable of representing a surface internally as a triangle mesh.

To display volumetric data as a surface, the triangles defining the border between body and background have to be generated. One of the most commonly applied algorithms recruted for this purpose is the marching cubes algoithm \cite{lorensen87}, which we also utilized here.

A virtual cube "marches" through the volumetric stack and inserts triangles, according to the intensity values at the vertices of the cube (see figure \ref{fig:surface}a).

The resulting list of triangles are displayed, taking into account attributes like color and transparency. For a smooth, realistic representation, vertex normals are calculated and used for shading. An example of a 3D surface is shown in figure \ref{fig:surface}c, figure \ref{fig:surface}b shows how the surface is composed of triangles.

\subsection{Volume rendering}
Another method to visualize a set of 2D slices is volume rendering.
The general idea is to put the 2D slices one behind another. To avoid that the top slice obscure slices below, each voxel (volume element) is assigned a transparency value, which is dependent on its intensity value: Brighter voxels are more opaque, darker ones are more transparent.
In Java 3D, this technique can be implemented using textures: Each slice has a rectangular geometry, which defines the border of its texture. The texture itself is given by the set of 2D slices. (A tutorial about volume rendering in Java 3D can be found at \cite{volrend}.)
Figure \ref{fig:volume}c shows a volume rendering. To display coloured stacks, the transparency value of each voxel is calculated from the mean intensity of the red, green and blue color component.

To separate different channels in a double staining, our software allows to select the color channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b. Additionally, it provides the functionality to edit volumes via virtual scissors (figure \ref{fig:volume}d).

\subsection{Orthoslices}
The example of an orthoslice representation can be seen in figure \ref{fig:orthoslices}.
It consists of 3 orthogonal slices through the displayed volume. Together with the rotation capabilities, scrolling the slices through the volume offers a great way to follow neuritic fibres in brain tissue.

\subsection{Properties common to surfaces, volume renderings and orthoslices}
In the following, each item of the 3D universe, be it an orthoslice, a surface or a volume rendering, is called a body.
While all three kinds of bodies are different in general, they have several properties in common.
First, there is the behaviour they show on user interaction. Bodies can be rotated and shifted by the use of the mouse: The user can select a body through a click on it and then drag or shift+drag it to rotate and translate it respectively. (Without selecting a body, the same action causes the whole universe to be rotated/translated).
Several other properties are common to surfaces, volume renderings and orthoslices:
\begin{itemize}
\item color: It is possible to either show a body in its original color (as in the original set of the 2D slices) or to adjust its red, green and blue color component arbitrarily.
\item threshold: In case of a surface, this is the previously mentioned isovalue of the calculated isosurface. In case of volume renderings and orthoslices, it defines the lowest voxel value displayed. Darker voxels are omitted from the 3D rendering.
\item transparency value
\item channels: The color channels of the original 2D slices, which should be used for displaying the 3D body (see figure \ref{fig:volume}a and \ref{fig:volume}b).
\end{itemize}

The user can either set these properties when adding a body to the 3D universe, or she can change existing properties lateron.

\subsection{Several bodies in the 3D universe}
For several reasons, it is desirable to show more than one body in the 3D world.
Figure \ref{fig:multiple} shows as an example a 3D world with an isosurface of a fly brain wholemount, together with the volume rendering of some coloured neuropiles (lobula, mushroom bodies and antennal lobes).

Having more than one body in the 3D world comes along with some difficulties for the implementation. To be useful, it must be possible to interactively transform each body individually, as well as to transform the 3D universe as a whole. As stated previously, the mechanism is implemented in the following way: If the user selects an individual body by clicking on it (the selection shows up by a red bounding box), only the selected body is transformed, the others keep their position and orientation. If no body is selected, the whole universe is transformed.
More details about how this is implemented in the software can be found later in the implementation details.

\subsection{Handling transformations}
There exist two ways to transform a body in the 3D universe. The first one, which has already been mentioned, is by direct user interaction via the mouse. Transformations are restricted to translation and rotation. (Scaling, or zooming, is possible as well, but not on the level of individual bodies).

Next to direct interactions via the mouse, transformations can be specified explicitly for each body, in the form of a 4x4 matrix. The user can set a transformation, she can concatenate the current transformation with a newly specified one, she can store the current transformation of the body to an external file and reload it.

This possibility to specify defined transformations on the level of individual bodies can for example be used to investigate and visualize the results of 3D image registration. Image registration is the transformation of a model image in such a way that it matches a reference image \cite{jenett2006}. Its purpose is for example the creation of an average staining out of images obtained from equally stained tissue of different individuals within one species. One loads the template image into the 3D world, without specifying a transformation, and subsequently the model image, with the transformation calculated by a registration algorithm. This enables one to manually judge the quality of the registration.

\subsection{Animation and movie recording}
Our software allows to animate the 3D universe. It is possible to record a movie of such an animation. The result is a stack of 2D images, which show subsequent frames of the animation (figure \ref{fig:animation}. This stack can be saved in video format, e.g. by ImageJ's avi exporting function. Video files like this one can be easily incorporated in powerpoint presentations.

\section{Implementation aspects}
\subsection*{Integration into ImageJ}
As mentioned previously, ImageJ consists of a tight core covering basic image processing functionality. This core can be extended individually by installing additional plugins, either developped on one's own or obtained from third parties.
To meet ImageJ's structure, the 3D viewer was also implemented as a plugin. It can easily be installed by downloading it and placing it into ImageJ's plugins directory.

\subsection*{Java 3D scene graph}
The scene graph is an arrangement of 3D objects in a tree structure that completely specifies the content of a virtual universe, and how it is to be rendered. It consists of nodes which specify the 3D objects by defining their geometry, appearance, location and transformation. There exist also nodes which define the properties of the universe like sight sources and sound. \cite{java3dtut}

Different types of nodes exist which have different function: The most prominent ones are:
\begin{itemize}
\item BranchGroup: A node that is capable of being the root of several subtrees.
\item TransformGroup: A Node that transforms the subtree below by a specified transformation.
\item Switch: A node having several child nodes, whose display can be toggled (i.e. which can be switched on and off).
\item Shape3D: A node representing a body in the 3D world. A Shape3D is specified by its Appearance and Geometry objects.
\item Geometry: Defines the geometry of the corresponding Shape3D, e.g. its vertex coordinates. In case of an isosurface, the geometry holds the triangle mesh, in the case of a volume rendering or orthoslices, it holds a set of rectangles.
\item Appearance: Defines several attributes of a Shape3D which influence its display, like color, transparency, textures etc.
\end{itemize}

Parts of the scene graph of the 3D viewer are shown in figure \ref{fig:scenegraph}. It particularly shows the path from the topmost root node to an attached Shape3D representing a surface. The structure is intended to elucidate the interplay of the different TransformGroups, which allows the rather complex transformation handling in the viewer. The nodes have the following meaning:

\begin{description}
\item[TG scale:] TransformGroup for global scaling. Acts either on user interaction (zooming) or automatically to fit a body into the 3D universe.
\item[TG rotate:] TransformGroup for global rotation. Acts on user interaction when no body is selected.
\item[TG translate:] TransformGroup for global translation. Acts on user interaction when no body is selected.
\item[TG center:] TransformGroup for centering displayed bodies in the universe. Each time updated when a new body is added to the universe.
\item[BG scene:] BranchGroup which holds a subtree for each displayed body.
\item[BG content:] Root BranchGroup of a body's subtree.
\item[TG l\_translate:]  TransformGroup for local translations. Acts on user interaction when a body is selected.
\item[TG l\_rotate:] TransformGroup for local rotation. Acts on user interaction when a body is selected.
\item[Switch bcSwitch:] Switch which holds next to the body's Shape3D a optionally displayable coordinate system and the bounding box which shows up on selection.
\item[Shape3D surface:] Shape3D which represents the surface.
\end{description}

\section{Availability}
The 3D viewer can be downloaded on our homepage at http://www.neurofly.de under Downloads $>$ ImageJ $>$ ImageJ 3D Viewer. On the same site, comprehensive documentation can be found, together with a screen movie demonstrating all the features of the viewer.

\bibliographystyle{plain}
\bibliography{bibliography}

\begin{figure}[m]
  \centering
  \includegraphics[width=\textwidth]{images/surface_montage.eps}
  \caption{Isosurface. Isosurfaces can be generated by the marching cubes algorithm. A cube is moved to the through the volume. The vertices of the cube are given by 8 neighbouring voxel values. Dependent on these 8 values, triangles are inserted at each position (a). The isosurface shown in magenta (c) was generated from a stack of 2D slices showing a fly brain. The images were obtained by confocal microscopy. In green one can see an optionally displayable coordinate system which indicates the origin of the original stack. In red one can see a bounding box which indicates that the brain is selected. In (c), the triangle mesh of the one of the optic lobes is shown.}
  \label{fig:surface}
\end{figure}

\begin{figure}[m]
\centering
\includegraphics[width=\textwidth]{images/volume_montage}
% \centering
% \subfloat[]{\includegraphics[width=0.45\textwidth]{images/volume_green}}
% \hspace{3mm}
% \subfloat[]{\includegraphics[width=0.45\textwidth]{images/volume_red}} \\
% \subfloat[]{\includegraphics[width=0.45\textwidth]{images/volume}}
% \hspace{3mm}
% \subfloat[]{\includegraphics[width=0.45\textwidth]{images/volume_edited}}
\caption {Volume rendering. This volume rendering shows also a fly brain. Alexa 488, a green dye, was used as a reference staining to colour the whole brain. Cy3, a red dye, was used to stain some specific brain regions. Where boths dyes coincide, the tissue shows up yellow. The viewer allows to specify the channels which should be displayed. (a) shows the green channel, (b) shows the red channel, (c) shows the red and green channel combined. The viewer allows additionally to edit volumes. (d) shows the same brain, with the right optic lobe cut off via the viewer's volume editing tool.}
\label{fig:volume}
\end{figure}

\begin{figure}[m]
\includegraphics[width=\textwidth]{images/orthoslices.eps}
\caption{Orthoslices. Orthoslices are 3 orthogonal slices through the volume. Shown here is the same fly brain as in figure \ref{fig:volume}. The slices are chosen in such a way that a section through the right peduncle, which was stained by the red and green dye, is visible.}
\label{fig:orthoslices}
\end{figure}

\begin{figure}[m]
\includegraphics[width=\textwidth]{images/multiple.eps}
\caption{Multiple bodies. The isosurface shown in figure \ref{fig:surface} is shown here in lightgray and a transparency value of 0.5. Additionally, the volume rendering of another stack is displayed, which shows lobula (orange), mushroom bodies (magenta) and antennal lobes (blue) of the same brain.}
\label{fig:multiple}
\end{figure}

\begin{figure}[m]
\includegraphics[width=\textwidth]{images/animation.eps}
\caption{Animation. The image sequence here shows a rotation of about 90 degree. An animation like this can be recorded and results in an image stack, which can be saved in a video format like AVI.}
\label{fig:animation}
\end{figure}

\begin{figure}[m]
\includegraphics[width=0.7\textwidth]{images/scenegraph.eps}
\caption{Part of the scenegraph of the 3D viewer. Shown here is an example path from the root node to a surface node. For more information about the meaning of the nodes see the text.}
\label{fig:scenegraph}
\end{figure}

\end{document}

### 2008-04-16 viewer_paper/paper.tex
Send paper.tex through aspell.

@@
 \begin{abstract}
-In this paper, we introduce a tool for visualizing 3D datasets.
+In this paper, we introduce a tool for visualising 3D data sets.
 Microscopy techniques in biology and medicine, like confocal microscopy, EM, MRI etc, often provide a stack of 2D images.
 For the investigation of these images, however, the two-dimensional view of the single slices is often not sufficient.
-We therefore developped a software for 3D visualisation, which allows to display the stacks optionally as surfaces, volume renderings or orthoslices.
-The software focusses on interactive, hardware-accelerated display of the stacks.
+We therefore developed a software for 3D visualisation, which allows to display the stacks optionally as surfaces, volume renderings or orthoslices.
+The software focuses on interactive, hardware-accelerated display of the stacks.
 Additionally, it offers possibilities for animation and movie recording, as well as an elaborate way to handle transformations of displayed bodies.
@@
 \item It is freely available and extensible.
-\item It was developped for simplicity of usage.
+\item It was developed for simplicity of usage.
 \item It is incorporated into ImageJ \cite{imagej}, a prominent image processing software for microscopy data.
@@
 \section{Introduction}
-Medical and biological imaging techniques often acquire 3D datasets in the form of stacks of 2D images.
+Medical and biological imaging techniques often acquire 3D data sets in the form of stacks of 2D images.
 Confocal microscopy for example can be used to cut a \textit{Drosophila} brain virtually into slices of a few $\mu$m thickness and scans it.
@@
 Its most relevant advantage however is its incorporation into ImageJ, which is one of the most wide-spread image processing packages in the field of microscopy.
-Users of ImageJ who already benefit from its numerous possibilities and plugins to process microscopic images, like filtering and deconvolution \cite{mbf}, can now visualize their data in 3D seemlessly, without the need of another software.
+Users of ImageJ who already benefit from its numerous possibilities and plugins to process microscopic images, like filtering and deconvolution \cite{mbf}, can now visualise their data in 3D seamlessly, without the need of another software.
 
-This paper is organized as follows: We first introduce briefly the technologies used for the implementation of the software.
+This paper is organised as follows: We first introduce briefly the technologies used for the implementation of the software.
 Subsequently, we show how the surface, volume and orthoslice views are created, as well as more information about the features of this software.
@@
 \textbf{ImageJ} \cite{imagej} is used as underlying platform.
-There are several reasons for choosing ImageJ for this task: Using an image processing platform in general has the advantage that the developper does not need to take care about low-level actions like opening/saving images etc, since functionality like that is given by the chosen platform.
+There are several reasons for choosing ImageJ for this task: Using an image processing platform in general has the advantage that the developer does not need to take care about low-level actions like opening/saving images etc, since functionality like that is given by the chosen platform.
 The choice fell on ImageJ since it is by far the most widely used image processing software in the biological and medical field \cite{mbf}.
@@
 \item It is freely available and open source.
-\item It consists of a compact core covering basic functionality, and is easily extendible to the user's needs by plugins.
+\item It consists of a compact core covering basic functionality, and is easily extensible to the user's needs by plugins.
 \item There exists a active and competent community, which both serves as support via online forums and as developers for plugins of which hundreds can be found on the internet.
@@
 \textbf{Java 3D.} To guarantee interactive 3D visualisation, hardware capabilities must be used optimally.
-Many 3D rendering tasks can nowadays be fullfilled by graphics cards, which frees the CPU and accelerates operations drastically.
+Many 3D rendering tasks can nowadays be fulfilled by graphics cards, which frees the CPU and accelerates operations drastically.
 Java 3D \cite{java3d} is a technology which allows to use these hardware capabilities from within Java.
@@
 \section{3D Visualisation}
-3D datasets can be represented in different ways. The options we implemented in this work are surfaces, volume renderings and orthoslices.
+3D data sets can be represented in different ways. The options we implemented in this work are surfaces, volume renderings and orthoslices.
 Starting point is a set of 2D slices which contain volumetric data, e.g.
@@
 A surface is generally understood as the border between a volumetric body and its environment.
-Given a volumetric 3D dataset, constructing a corresponding surface is an ill-defined problem.
+Given a volumetric 3D data set, constructing a corresponding surface is an ill-defined problem.
 In computer graphics, one often uses isosurfaces for this purpose.
@@
 To display volumetric data as a surface, the triangles defining the border between body and background have to be generated.
-One of the most commonly applied algorithms recruted for this purpose is the marching cubes algoithm \cite{lorensen87}, which we also utilized here.
+One of the most commonly applied algorithms recruited for this purpose is the marching cubes algorithm \cite{lorensen87}, which we also utilised here.
 
-A virtual cube "marches" through the volumetric stack and inserts triangles, according to the intensity values at the vertices of the cube (see figure \ref{fig:surface}a).
+A virtual cube "marches" through the volumetric stack and inserts triangles, according to the intensity values at the vertexes of the cube (see figure \ref{fig:surface}a).
 
-The resulting list of triangles are displayed, taking into account attributes like color and transparency.
+The resulting list of triangles are displayed, taking into account attributes like colour and transparency.
 For a smooth, realistic representation, vertex normals are calculated and used for shading.
@@
 \subsection{Volume rendering}
-Another method to visualize a set of 2D slices is volume rendering.
+Another method to visualise a set of 2D slices is volume rendering.
 The general idea is to put the 2D slices one behind another.
@@
 (A tutorial about volume rendering in Java 3D can be found at \cite{volrend}.)
-Figure \ref{fig:volume}c shows a volume rendering. To display coloured stacks, the transparency value of each voxel is calculated from the mean intensity of the red, green and blue color component.
+Figure \ref{fig:volume}c shows a volume rendering. To display coloured stacks, the transparency value of each voxel is calculated from the mean intensity of the red, green and blue colour component.
 
-To separate different channels in a double staining, our software allows to select the color channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b.
+To separate different channels in a double staining, our software allows to select the colour channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b.
 Additionally, it provides the functionality to edit volumes via virtual scissors (figure \ref{fig:volume}d).
@@
 \begin{itemize}
-\item color: It is possible to either show a body in its original color (as in the original set of the 2D slices) or to adjust its red, green and blue color component arbitrarily.
+\item colour: It is possible to either show a body in its original colour (as in the original set of the 2D slices) or to adjust its red, green and blue colour component arbitrarily.
 \item threshold: In case of a surface, this is the previously mentioned isovalue of the calculated isosurface.
@@
 \item transparency value
-\item channels: The color channels of the original 2D slices, which should be used for displaying the 3D body (see figure \ref{fig:volume}a and \ref{fig:volume}b).
+\item channels: The colour channels of the original 2D slices, which should be used for displaying the 3D body (see figure \ref{fig:volume}a and \ref{fig:volume}b).
 \end{itemize}
 
-The user can either set these properties when adding a body to the 3D universe, or she can change existing properties lateron.
+The user can either set these properties when adding a body to the 3D universe, or she can change existing properties later on.
 
@@
 For several reasons, it is desirable to show more than one body in the 3D world.
-Figure \ref{fig:multiple} shows as an example a 3D world with an isosurface of a fly brain wholemount, together with the volume rendering of some coloured neuropiles (lobula, mushroom bodies and antennal lobes).
+Figure \ref{fig:multiple} shows as an example a 3D world with an isosurface of a fly brain whole-mount, together with the volume rendering of some coloured neuropiles (lobula, mushroom bodies and antennal lobes).
 
@@
 
-This possibility to specify defined transformations on the level of individual bodies can for example be used to investigate and visualize the results of 3D image registration.
+This possibility to specify defined transformations on the level of individual bodies can for example be used to investigate and visualise the results of 3D image registration.
 Image registration is the transformation of a model image in such a way that it matches a reference image \cite{jenett2006}.
@@
 by ImageJ's avi exporting function.
-Video files like this one can be easily incorporated in powerpoint presentations.
+Video files like this one can be easily incorporated in PowerPoint presentations.
 
@@
 As mentioned previously, ImageJ consists of a tight core covering basic image processing functionality.
-This core can be extended individually by installing additional plugins, either developped on one's own or obtained from third parties.
+This core can be extended individually by installing additional plugins, either developed on one's own or obtained from third parties.
 To meet ImageJ's structure, the 3D viewer was also implemented as a plugin. It can easily be installed by downloading it and placing it into ImageJ's plugins directory.
@@
 In case of an isosurface, the geometry holds the triangle mesh, in the case of a volume rendering or orthoslices, it holds a set of rectangles.
-\item Appearance: Defines several attributes of a Shape3D which influence its display, like color, transparency, textures etc.
+\item Appearance: Defines several attributes of a Shape3D which influence its display, like colour, transparency, textures etc.
 \end{itemize}
@@
 \item[TG translate:] TransformGroup for global translation. Acts on user interaction when no body is selected.
-\item[TG center:] TransformGroup for centering displayed bodies in the universe. Each time updated when a new body is added to the universe.
+\item[TG center:] TransformGroup for centring displayed bodies in the universe. Each time updated when a new body is added to the universe.
 \item[BG scene:] BranchGroup which holds a subtree for each displayed body.
@@
 A cube is moved to the through the volume.
-The vertices of the cube are given by 8 neighbouring voxel values.
+The vertexes of the cube are given by 8 neighbouring voxel values.
 Dependent on these 8 values, triangles are inserted at each position (a).
@@
 Cy3, a red dye, was used to stain some specific brain regions.
-Where boths dyes coincide, the tissue shows up yellow.
+Where both dyes coincide, the tissue shows up yellow.
 The viewer allows to specify the channels which should be displayed.
@@
 \caption{Multiple bodies.
-The isosurface shown in figure \ref{fig:surface} is shown here in lightgray and a transparency value of 0.5.
+The isosurface shown in figure \ref{fig:surface} is shown here in light grey and a transparency value of 0.5.
 Additionally, the volume rendering of another stack is displayed, which shows lobula (orange), mushroom bodies (magenta) and antennal lobes (blue) of the same brain.}

### 2008-04-16 viewer_paper/paper.tex
Still 2 typos...

@@
 The general idea is to put the 2D slices one behind another.
-To avoid that the top slice obscure slices below, each voxel (volume element) is assigned a transparency value, which is dependent on its intensity value: Brighter voxels are more opaque, darker ones are more transparent.
+To avoid that the top slice obscures slices below, each voxel (volume element) is assigned a transparency value, which is dependent on its intensity value: Brighter voxels are more opaque, darker ones are more transparent.
 In Java 3D, this technique can be implemented using textures: Each slice has a rectangular geometry, which defines the border of its texture.
@@
 
-To separate different channels in a double staining, our software allows to select the colour channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b.
+To separate different channels in a double staining, our software allows to select the colour channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b).
 Additionally, it provides the functionality to edit volumes via virtual scissors (figure \ref{fig:volume}d).

### 2008-04-16 viewer_paper/paper.tex
paper: One more word to orthoslices.

@@
 \subsection{Orthoslices}
-The example of an orthoslice representation can be seen in figure \ref{fig:orthoslices}.
-It consists of 3 orthogonal slices through the displayed volume.
+The example for an orthoslice representation can be seen in figure \ref{fig:orthoslices}.
+It is composed of 3 orthogonal slices through the volume, where each slice is parallel to two coordinate axes.
+This is opposed to oblique slices, which display arbitrary sections through a volume.
+
 Together with the rotation capabilities, scrolling the slices through the volume offers a great way to follow neuritic fibres in brain tissue.

### 2008-04-17 ij3d/Image3DMenubar.java
In the 3D viewer, allow to select a content via the menu. A new MenuItem is created which displays all the contents of the universe and selects them on click.

@@ -9,6 +9,7 @@ import math3d.Transform_IO;
 import java.awt.event.*;
 import java.awt.*;
 import java.util.Vector;
+import java.util.Iterator;
 
 import orthoslice.OrthoGroup;
 import voltex.VoltexGroup;
@@ -60,10 +61,13 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 	private CheckboxMenuItem lock;
 
 	private Menu selectedMenu;
+	private Menu selectSubMenu;
 	private Menu viewMenu;
 	private Menu contentsMenu;
 	private Menu fileMenu;
 
+	// These strings are the names of the stataic methods in
+	// ImageJ3DViewer.
 	public static final String START_ANIMATE = "startAnimate";
 	public static final String STOP_ANIMATE = "stopAnimate";
 	public static final String START_RECORD = "startRecord";
@@ -93,6 +97,8 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 	public static final String DELETE = "delete";
 
 	public static final String SMOOTH = "smooth";
+
+	private SelectionListener selListener = new SelectionListener();
 
 	public Image3DMenubar(Image3DUniverse univ) {
 		super();
@@ -193,6 +199,11 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 
 		universe.addSeparator();
 
+		selectSubMenu = createSelectSubMenu();
+		universe.add(selectSubMenu);
+
+		universe.addSeparator();
+
 		delete = new MenuItem("Delete");
 		delete.setEnabled(false);
 		delete.addActionListener(this);
@@ -205,6 +216,18 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 		universe.add(smoothAllMeshes);
 
 		return universe;
+	}
+
+	public Menu createSelectSubMenu() {
+		Menu select = new Menu("Select");
+		if(univ == null)
+			return select;
+		for(Iterator it = univ.contents(); it.hasNext();) {
+			String name = ((Content)it.next()).getName();
+			MenuItem mi = new MenuItem(name);
+			mi.addActionListener(selListener);
+		}
+		return select;
 	}
 
 	public Menu createSelectedMenu() {
@@ -269,6 +292,15 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 		content.add(saveTransform);
 
 		return content;
+	}
+
+	private class SelectionListener implements ActionListener {
[…]

### 2008-04-17 isosurface/IsoShape.java
For each Content in the 3D viewer, calculate minPoint, maxPoint and centerPoint according to pixel values: minPoint is the minimal non-zero coordinate maxPoint is the maximal non-zero coordinate centerPoint is the center of gravity
This has 2 effects:
* It shrinks the red bounding boxes to the actual volume, discarding zero voxels
* It makes local rotations rotating around the cog, which is what one
  intuitively expects.

@@ -69,8 +69,10 @@ public final class IsoShape extends Shape3D {
 		if(mesh == null) return;
 		min.x = min.y = min.z = Float.MAX_VALUE;
 		max.x = max.y = max.z = Float.MIN_VALUE;
+		center.x = center.y = center.z = 0;
 		
-		for(int i = 0; i < mesh.size(); i++) {
+		int n = mesh.size();
+		for(int i = 0; i < n; i++) {
 			Point3f p = (Point3f)mesh.get(i);
 			if(p.x < min.x) min.x = p.x;
 			if(p.y < min.y) min.y = p.y;
@@ -78,10 +80,13 @@ public final class IsoShape extends Shape3D {
 			if(p.x > max.x) max.x = p.x;
 			if(p.y > max.y) max.y = p.y;
 			if(p.z > max.z) max.z = p.z;
+			center.x += p.x;
+			center.y += p.y;
+			center.z += p.z;
 		}
-		center.x = (max.x-min.x)/2;
-		center.y = (max.y-min.y)/2;
-		center.z = (max.z-min.z/2);
+		center.x /= n;
+		center.y /= n;
+		center.z /= n;
 	}
 
 	public void setColor(Color3f color) {

### 2008-04-17 orthoslice/OrthoGroup.java

@@ -63,16 +63,59 @@ public class OrthoGroup extends Content {
 
 	public void calculateMinMaxCenterPoint() {
 		ImagePlus imp = getImage();
+		int w = imp.getWidth(), h = imp.getHeight();
+		int d = imp.getStackSize();
 		Calibration c = imp.getCalibration();
 		minPoint = new Point3f();
-		maxPoint = new Point3f((float)(imp.getWidth()*c.pixelWidth),
-				(float)(imp.getHeight()*c.pixelHeight),
-				(float)(imp.getStackSize()*c.pixelDepth));
-		centerPoint = new Point3f(maxPoint.x/2, maxPoint.y/2, 
-				maxPoint.z/2);
+		maxPoint = new Point3f();
+		centerPoint = new Point3f();
+		minPoint.x = w * (float)c.pixelHeight;
+		minPoint.y = h * (float)c.pixelHeight;
+		minPoint.z = d * (float)c.pixelDepth;
+		maxPoint.x = 0;
+		maxPoint.y = 0;
+		maxPoint.z = 0;
+
+		long count = 0;
+		for(int zi = 0; zi < d; zi++) {
+			float z = zi * (float)c.pixelDepth;
+			byte[] p = (byte[])imp.getStack().getPixels(zi+1);
+			for(int i = 0; i < p.length; i++) {
+				if(p[i] == 0) continue;
+				float x = (i % w) * (float)c.pixelWidth;
+				float y = (i / w) * (float)c.pixelHeight;
+				if(x < minPoint.x) minPoint.x = x;
+				if(y < minPoint.y) minPoint.y = y;
+				if(z < minPoint.z) minPoint.z = z;
+				if(x > maxPoint.x) maxPoint.x = x;
+				if(y > maxPoint.y) maxPoint.y = y;
+				if(z > maxPoint.z) maxPoint.z = z;
+				centerPoint.x += x;
+				centerPoint.y += y;
+				centerPoint.z += z;
+				count++;
+			}
+		}
+		centerPoint.x /= count;
+		centerPoint.y /= count;
+		centerPoint.z /= count;
+		
 		createBoundingBox();
 		showBoundingBox(false);
 	}
+
+// 	public void calculateMinMaxCenterPoint() {
+// 		ImagePlus imp = getImage();
+// 		Calibration c = imp.getCalibration();
+// 		minPoint = new Point3f();
[…]

### 2008-04-17 voltex/VoltexGroup.java

@@ -63,16 +63,59 @@ public class VoltexGroup extends Content {
 
 	public void calculateMinMaxCenterPoint() {
 		ImagePlus imp = getImage();
+		int w = imp.getWidth(), h = imp.getHeight();
+		int d = imp.getStackSize();
 		Calibration c = imp.getCalibration();
 		minPoint = new Point3f();
-		maxPoint = new Point3f((float)(imp.getWidth()*c.pixelWidth),
-				(float)(imp.getHeight()*c.pixelHeight),
-				(float)(imp.getStackSize()*c.pixelDepth));
-		centerPoint = new Point3f(maxPoint.x/2, maxPoint.y/2, 
-				maxPoint.z/2);
+		maxPoint = new Point3f();
+		centerPoint = new Point3f();
+		minPoint.x = w * (float)c.pixelHeight;
+		minPoint.y = h * (float)c.pixelHeight;
+		minPoint.z = d * (float)c.pixelDepth;
+		maxPoint.x = 0;
+		maxPoint.y = 0;
+		maxPoint.z = 0;
+
+		long count = 0;
+		for(int zi = 0; zi < d; zi++) {
+			float z = zi * (float)c.pixelDepth;
+			byte[] p = (byte[])imp.getStack().getPixels(zi+1);
+			for(int i = 0; i < p.length; i++) {
+				if(p[i] == 0) continue;
+				float x = (i % w) * (float)c.pixelWidth;
+				float y = (i / w) * (float)c.pixelHeight;
+				if(x < minPoint.x) minPoint.x = x;
+				if(y < minPoint.y) minPoint.y = y;
+				if(z < minPoint.z) minPoint.z = z;
+				if(x > maxPoint.x) maxPoint.x = x;
+				if(y > maxPoint.y) maxPoint.y = y;
+				if(z > maxPoint.z) maxPoint.z = z;
+				centerPoint.x += x;
+				centerPoint.y += y;
+				centerPoint.z += z;
+				count++;
+			}
+		}
+		centerPoint.x /= count;
+		centerPoint.y /= count;
+		centerPoint.z /= count;
+		
 		createBoundingBox();
 		showBoundingBox(false);
 	}
+
+// 	public void calculateMinMaxCenterPoint() {
+// 		ImagePlus imp = getImage();
+// 		Calibration c = imp.getCalibration();
+// 		minPoint = new Point3f();
[…]

### 2008-04-18 ij3d/Image3DMenubar.java
New feature for the viewer:
Display properties for a content, like center of gravity and volume. The
latter one is particularly interesting for surfaces. (see MeshProperties.java).

@@ -3,6 +3,7 @@ package ij3d;
 import ij.gui.GenericDialog;
 import ij.IJ;
 import ij.ImagePlus;
+import ij.text.TextWindow;
 
 import math3d.Transform_IO;
 
@@ -40,6 +41,7 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 	private MenuItem fill;
 	private MenuItem slices;
 	private MenuItem delete;
+	private MenuItem properties;
 	private MenuItem windowSize;
 	private MenuItem resetView;
 	private MenuItem startRecord;
@@ -271,6 +273,12 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 
 		content.addSeparator();
 
+		properties = new MenuItem("Properties");
+		properties.addActionListener(this);
+		content.add(properties);
+
+		content.addSeparator();
+
 		lock = new CheckboxMenuItem("Lock");
 		lock.addItemListener(this);
 		content.add(lock);
@@ -488,6 +496,27 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 				univ.fireTransformationFinished();
 			}
 			record(SET_TRANSFORM, affine2string(t));
+		}
+
+		if(e.getSource() == properties) {
+			Content c = univ.getSelected();
+			if(c == null) {
+				IJ.error("Selection required");
+				return;
+			}
+			TextWindow tw = new TextWindow(c.getName(), 
+				" \tx\ty\tz",
+				"min\t" + (float)c.minPoint.x + "\t"
+					+ (float)c.minPoint.y + "\t"
+					+ (float)c.minPoint.z + "\n" +
+				"max\t" + (float)c.maxPoint.x + "\t"
+					+ (float)c.maxPoint.y + "\t"
+					+ (float)c.maxPoint.z + "\n" +
+				"cog\t" + (float)c.centerPoint.x + "\t"
+					+ (float)c.centerPoint.y + "\t"
+					+ (float)c.centerPoint.z + "\n\n" +
+				"volume\t" + c.getVolume(),
+				512, 512);
 		}
 
 		if(e.getSource() == applyTransform) {

### 2008-04-18 isosurface/IsoShape.java

@@ -20,6 +20,7 @@ public final class IsoShape extends Shape3D {
 	List mesh = null;
 	private int threshold;
 	private float transparency;
+	private float volume;
 
 	public IsoShape(List mesh, int threshold, float transparency){
 		this.mesh = mesh;
@@ -30,6 +31,7 @@ public final class IsoShape extends Shape3D {
 		this.setCapability(ALLOW_APPEARANCE_READ);
 		this.setCapability(ALLOW_APPEARANCE_WRITE);
 		this.update();
+
 	}
 
 	public IsoShape(List mesh, Color3f color, int threshold, 
@@ -66,13 +68,13 @@ public final class IsoShape extends Shape3D {
 
 	public void calculateMinMaxCenterPoint(Point3f min, 
 				Point3f max, Point3f center) {
-		if(mesh == null) return;
+
+		if(mesh == null)
+			return;
+
 		min.x = min.y = min.z = Float.MAX_VALUE;
 		max.x = max.y = max.z = Float.MIN_VALUE;
-		center.x = center.y = center.z = 0;
-		
-		int n = mesh.size();
-		for(int i = 0; i < n; i++) {
+		for(int i = 0; i < mesh.size(); i++) {
 			Point3f p = (Point3f)mesh.get(i);
 			if(p.x < min.x) min.x = p.x;
 			if(p.y < min.y) min.y = p.y;
@@ -80,13 +82,13 @@ public final class IsoShape extends Shape3D {
 			if(p.x > max.x) max.x = p.x;
 			if(p.y > max.y) max.y = p.y;
 			if(p.z > max.z) max.z = p.z;
-			center.x += p.x;
-			center.y += p.y;
-			center.z += p.z;
 		}
-		center.x /= n;
-		center.y /= n;
-		center.z /= n;
+		double[][] inertia = new double[3][3];
+		volume = (float)MeshProperties.compute(mesh, center, inertia);
+	}
+
+	public float getVolume() {
+		return volume;
 	}
 
 	public void setColor(Color3f color) {

### 2008-04-18 isosurface/MeshGroup.java

@@ -104,6 +104,12 @@ public class MeshGroup extends Content {
 		}
 	}
 
+	public float getVolume() {
+		if(shape == null)
+			return -1;
+		return shape.getVolume();
+	}
+
 	public void colorUpdated(Color3f oldColor, Color3f newColor) {
 		if(newColor == null){
 			int val = getImage().getProcessor().

### 2008-04-18 orthoslice/OrthoGroup.java

@@ -37,6 +37,7 @@ public class OrthoGroup extends Content {
 	private Renderer renderer;
 	private TransformGroup tg;
 	private int[] slices;
+	private float volume;
 
 	public OrthoGroup(String name, Color3f color, ImagePlus image, 
 		boolean[] channels, int resamplingF) {
@@ -76,12 +77,13 @@ public class OrthoGroup extends Content {
 		maxPoint.y = 0;
 		maxPoint.z = 0;
 
-		long count = 0;
+		long vol = 0;
 		for(int zi = 0; zi < d; zi++) {
 			float z = zi * (float)c.pixelDepth;
 			byte[] p = (byte[])imp.getStack().getPixels(zi+1);
 			for(int i = 0; i < p.length; i++) {
 				if(p[i] == 0) continue;
+				vol += (p[i] & 0xff);
 				float x = (i % w) * (float)c.pixelWidth;
 				float y = (i / w) * (float)c.pixelHeight;
 				if(x < minPoint.x) minPoint.x = x;
@@ -90,15 +92,18 @@ public class OrthoGroup extends Content {
 				if(x > maxPoint.x) maxPoint.x = x;
 				if(y > maxPoint.y) maxPoint.y = y;
 				if(z > maxPoint.z) maxPoint.z = z;
-				centerPoint.x += x;
-				centerPoint.y += y;
-				centerPoint.z += z;
-				count++;
+				centerPoint.x += (p[i] & 0xff) * x;
+				centerPoint.y += (p[i] & 0xff) * y;
+				centerPoint.z += (p[i] & 0xff) * z;
 			}
 		}
-		centerPoint.x /= count;
-		centerPoint.y /= count;
-		centerPoint.z /= count;
+		centerPoint.x /= vol;
+		centerPoint.y /= vol;
+		centerPoint.z /= vol;
+
+		volume = (float)(vol * c.pixelWidth
+				* c.pixelHeight
+				* c.pixelDepth);
 		
 		createBoundingBox();
 		showBoundingBox(false);
@@ -116,6 +121,10 @@ public class OrthoGroup extends Content {
 // 		createBoundingBox();
 // 		showBoundingBox(false);
 // 	}
+
+	public float getVolume() {
+		return volume;
+	}
 		
 	public static OrthoGroup addContent(Image3DUniverse univ, 
 							ImagePlus grey) {

### 2008-04-18 voltex/VoltexGroup.java

@@ -35,6 +35,7 @@ public class VoltexGroup extends Content {
 
 	private Renderer renderer;
 	private TransformGroup tg;
+	private float volume;
 
 	public VoltexGroup(String name, Color3f color, ImagePlus image, 
 		boolean[] channels, int resamplingF) {
@@ -76,12 +77,13 @@ public class VoltexGroup extends Content {
 		maxPoint.y = 0;
 		maxPoint.z = 0;
 
-		long count = 0;
+		long vol = 0;
 		for(int zi = 0; zi < d; zi++) {
 			float z = zi * (float)c.pixelDepth;
 			byte[] p = (byte[])imp.getStack().getPixels(zi+1);
 			for(int i = 0; i < p.length; i++) {
 				if(p[i] == 0) continue;
+				vol += (p[i] & 0xff);
 				float x = (i % w) * (float)c.pixelWidth;
 				float y = (i / w) * (float)c.pixelHeight;
 				if(x < minPoint.x) minPoint.x = x;
@@ -90,18 +92,25 @@ public class VoltexGroup extends Content {
 				if(x > maxPoint.x) maxPoint.x = x;
 				if(y > maxPoint.y) maxPoint.y = y;
 				if(z > maxPoint.z) maxPoint.z = z;
-				centerPoint.x += x;
-				centerPoint.y += y;
-				centerPoint.z += z;
-				count++;
+				centerPoint.x += (p[i] & 0xff) * x;
+				centerPoint.y += (p[i] & 0xff) * y;
+				centerPoint.z += (p[i] & 0xff) * z;
 			}
 		}
-		centerPoint.x /= count;
-		centerPoint.y /= count;
-		centerPoint.z /= count;
+		centerPoint.x /= vol;
+		centerPoint.y /= vol;
+		centerPoint.z /= vol;
+
+		volume = (float)(vol * c.pixelWidth
+				* c.pixelHeight
+				* c.pixelDepth);
 		
 		createBoundingBox();
 		showBoundingBox(false);
+	}
+
+	public float getVolume() {
+		return volume;
 	}
 
 // 	public void calculateMinMaxCenterPoint() {

### 2008-05-04 tracing/AutoPoint.java
Vaguely working version

@@ -26,18 +26,49 @@ public class AutoPoint {
 	public int y;
 	public int z;
 	public boolean overThreshold = false;
+	public AutoPoint [] predecessors;
 	public AutoPoint(int x,int y,int z) {
 		this.x = x;
 		this.y = y;
 		this.z = z;
+		this.predecessors = null;
 	}
 	@Override
-		public String toString() {
+	public String toString() {
 		return "("+x+","+y+","+z+")";
 	}
 	@Override
-		public boolean equals(Object o) {
+	public boolean equals(Object o) {
 		AutoPoint op=(AutoPoint)o;
-		return (this.x == op.x) && (this.y == op.y) && (this.z == op.z);
+		// System.out.println("Testing equality between "+this+" and "+op);
+		boolean result = (this.x == op.x) && (this.y == op.y) && (this.z == op.z);
+		return result;
+	}
+	public void addPredecessor(AutoPoint p) {
+		if( predecessors == null ) {
+			predecessors = new AutoPoint[1];
+			predecessors[0] = p;
+		} else {
+			for( int i = 0; i < predecessors.length; ++i )
+				if( p.equals( predecessors[i] ) )
+					return;
+			AutoPoint [] n = new AutoPoint[predecessors.length+1];
+			System.arraycopy(predecessors,0,n,0,predecessors.length);
+			n[predecessors.length] = p;
+			predecessors = n;
+		}
+	}
+	public void addPredecessors(AutoPoint [] newPredecessors) {
+		if( newPredecessors == null )
+			return;
+		if( predecessors == null ) {
+			predecessors = new AutoPoint[newPredecessors.length];
+			System.arraycopy(newPredecessors,0,predecessors,0,newPredecessors.length);
+		} else {
+			AutoPoint [] n = new AutoPoint[predecessors.length+newPredecessors.length];
+			System.arraycopy(predecessors,0,n,0,predecessors.length);
+			System.arraycopy(newPredecessors,0,n,predecessors.length,newPredecessors.length);
+			predecessors = n;
+		}
 	}
 }

### 2008-05-04 tracing/AutoSearchThread.java

@@ -33,10 +33,13 @@ public class AutoSearchThread extends SearchThread {
 	float [][] tubeValues;
 	float tubenessThreshold;
 
-	SinglePathsGraph thisPathGraph;
 	SinglePathsGraph previousPathGraph;
 
 	ArrayList<AutoPoint> destinations = new ArrayList(512);
+
+	public ArrayList<AutoPoint> getDestinations() {
+		return destinations;
+	}
 
 	int start_x, start_y, start_z;
 
@@ -60,7 +63,6 @@ public class AutoSearchThread extends SearchThread {
 		this.tubenessThreshold = tubenessThreshold;
 
 		this.previousPathGraph = previousPathGraph;
-		this.thisPathGraph = new SinglePathsGraph(width,height,depth);
 
 		this.start_x = startPoint.x;
 		this.start_y = startPoint.y;
@@ -106,5 +108,8 @@ public class AutoSearchThread extends SearchThread {
 		return 0;
         }
 
+	Path getPathBack( int from_x, int from_y, int from_z ) {
+		return nodes_as_image[from_z][from_y*width+from_x].asPath();
+	}
 
 }

### 2008-05-04 tracing/Auto_Tracer.java

@@ -31,6 +31,8 @@ import ij.plugin.PlugIn;
 import java.io.File;
 import java.util.PriorityQueue;
 import java.util.Comparator;
+import java.util.ArrayList;
+import java.util.Iterator;
 import util.BatchOpener;
 
 import java.awt.Color;
@@ -141,14 +143,16 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 		while( mostTubelikePoints.size() > 0 ) {
 
-			// Now get the most tubelike point:
-			AutoPoint p=mostTubelikePoints.poll();
+			System.out.println("=== Priority queue now has: "+mostTubelikePoints.size());
 
-			System.out.println("Got point "+p);
+			// Now get the most tubelike point:
+			AutoPoint startPoint=mostTubelikePoints.poll();
+
+			System.out.println("Got point "+startPoint);
 			
 			ast = new AutoSearchThread( image, /* original image */
 						    tubeValues, /* the "tubeness" filtered image */
-						    p, /* the point to start the search from */
+						    startPoint, /* the point to start the search from */
 						    tubenessThreshold,
 						    completePaths );
 
@@ -167,23 +171,115 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				ast.join();
 			} catch( InterruptedException e ) { }
 
-			
-			SinglePathsGraph recentGraph = ast.thisPathGraph;
-
-			/* 
-			   Now start the pruning...
-
-			 */
-
-			// Now remove all the points genuinely found in this search:
-
-			
-
-			// And merge the new graph into the ones so far:
+			boolean verbose = false;
 		      
+			// Now start the pruning:
+
+			ArrayList<AutoPoint> destinations = ast.getDestinations();
+			if ( verbose ) System.out.print("  === Destinations: "+destinations.size()+" ");
+			if ( verbose ) System.out.flush();
 			
+			int [] pa = new int[3];
 
+			for( Iterator<AutoPoint> it = destinations.iterator(); it.hasNext(); ) {
+				
+				if ( verbose ) System.out.print("    ");
 
+				AutoPoint d = it.next();
+				Path path = ast.getPathBack(d.x,d.y,d.z);
+
+				float [] rollingTubeness = new float[rollingLength];
+				int nextRollingAt = 0;
+				int slotsFilled = 0;
+
+				int lastIndex = path.size() - 1;
+				
[…]

### 2008-05-04 tracing/SinglePathsGraph.java

@@ -24,21 +24,32 @@ package tracing;
 import java.util.Hashtable;
 
 public class SinglePathsGraph {
-
-    int width, height, depth;
-
-    public SinglePathsGraph( int width, int height, int depth ) {
-        this.width = width;
-        this.height = height;
-        this.depth = depth;
-    }
-
-    // For fast lookup from positions:
-    Hashtable<Integer,AutoPoint> fromPosition=new Hashtable<Integer,AutoPoint>();
-
-    public AutoPoint get( int x, int y, int z ) {
-        int k = x + y * width + z * width * height;
-        return fromPosition.get(k);
-    }
-
+	
+	int width, height, depth;
+	
+	public SinglePathsGraph( int width, int height, int depth ) {
+		this.width = width;
+		this.height = height;
+		this.depth = depth;
+	}
+	
+	// For fast lookup from positions:
+	Hashtable<Integer,AutoPoint> fromPosition=new Hashtable<Integer,AutoPoint>();
+	
+	public AutoPoint get( int x, int y, int z ) {
+		int k = x + y * width + z * width * height;
+		return fromPosition.get(k);
+	}
+	
+	public void addPoint( AutoPoint p ) {
+		int k = p.x + p.y * width + p.z * width * height;
+		AutoPoint existingPoint=fromPosition.get( k );
+		if( existingPoint == null ) {
+			fromPosition.put(k,p);
+		} else {
+			// "merge" this point with the exisiting one -
+			// i.e. just add the predecessors:
+			existingPoint.addPredecessors(p.predecessors);
+		}
+	}	
 }

### 2008-05-05 tracing/InteractiveTracerCanvas.java
Double buffer in TracerCanvas instead of the subclasses

@@ -123,47 +123,6 @@ public class InteractiveTracerCanvas extends TracerCanvas implements KeyListener
 	
 	public void keyTyped(KeyEvent e) {}	
 	
-	/* Keep another Graphics for double-buffering... */
-	
-	private int backBufferWidth;
-	private int backBufferHeight;
-	
-	private Graphics backBufferGraphics;
-	private Image backBufferImage;
-	
-	private void resetBackBuffer() {
-		
-		if(backBufferGraphics!=null){
-			backBufferGraphics.dispose();
-			backBufferGraphics=null;
-		}
-		
-		if(backBufferImage!=null){
-			backBufferImage.flush();
-			backBufferImage=null;
-		}
-		
-		backBufferWidth=getSize().width;
-		backBufferHeight=getSize().height;
-		
-		backBufferImage=createImage(backBufferWidth,backBufferHeight);
-	        backBufferGraphics=backBufferImage.getGraphics();
-	}
-	
-	@Override
-	public void paint(Graphics g) {
-		
-		if(backBufferWidth!=getSize().width ||
-		   backBufferHeight!=getSize().height ||
-		   backBufferImage==null ||
-		   backBufferGraphics==null)
-			resetBackBuffer();
-		
-		super.paint(backBufferGraphics);
-		drawOverlay(backBufferGraphics);
-		g.drawImage(backBufferImage,0,0,this);
-	}
-	
 	@Override
 	public void mouseMoved( MouseEvent e ) {
 		

### 2008-05-05 tracing/TracerCanvas.java

@@ -156,5 +156,45 @@ public class TracerCanvas extends ThreePanesCanvas {
 
 	}
 
+	/* Keep another Graphics for double-buffering... */
+	
+	private int backBufferWidth;
+	private int backBufferHeight;
+	
+	private Graphics backBufferGraphics;
+	private Image backBufferImage;
+	
+	private void resetBackBuffer() {
+		
+		if(backBufferGraphics!=null){
+			backBufferGraphics.dispose();
+			backBufferGraphics=null;
+		}
+		
+		if(backBufferImage!=null){
+			backBufferImage.flush();
+			backBufferImage=null;
+		}
+		
+		backBufferWidth=getSize().width;
+		backBufferHeight=getSize().height;
+		
+		backBufferImage=createImage(backBufferWidth,backBufferHeight);
+	        backBufferGraphics=backBufferImage.getGraphics();
+	}
+	
+	@Override
+	public void paint(Graphics g) {
+		
+		if(backBufferWidth!=getSize().width ||
+		   backBufferHeight!=getSize().height ||
+		   backBufferImage==null ||
+		   backBufferGraphics==null)
+			resetBackBuffer();
+		
+		super.paint(backBufferGraphics);
+		drawOverlay(backBufferGraphics);
+		g.drawImage(backBufferImage,0,0,this);
+	}
 
 }

### 2008-05-05 tracing/Auto_Tracer.java
Added a "maxLoops" variable for testing purposes

@@ -86,6 +86,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 	}
 
+	boolean verbose = false;
+
 	public void autoTrace( ImagePlus image ) {
 
 		FileInfo originalFileInfo = image.getOriginalFileInfo();
@@ -141,8 +143,14 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 		SinglePathsGraph completePaths = new SinglePathsGraph(width,height,depth);
 
+		int maxLoops = 4;
+		int loopsDone = 0;
+
 		while( mostTubelikePoints.size() > 0 ) {
 
+			if( maxLoops >= 0 && loopsDone > maxLoops )
+				break;
+			
 			System.out.println("=== Priority queue now has: "+mostTubelikePoints.size());
 
 			// Now get the most tubelike point:
@@ -171,8 +179,6 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				ast.join();
 			} catch( InterruptedException e ) { }
 
-			boolean verbose = false;
-		      
 			// Now start the pruning:
 
 			ArrayList<AutoPoint> destinations = ast.getDestinations();
@@ -278,9 +284,12 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 					}
 				}
 
-				System.out.println("");
+				if (verbose) System.out.println("");
 			}
+		
+			++loopsDone;
 		}
+
 	}
 
 	AutoSearchThread ast;
@@ -407,7 +416,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		long currentTime = System.currentTimeMillis();
 		long timeSinceStarted = currentTime - timeStarted;
 		if( (inOpen + inClosed) > maxNodes || (timeSinceStarted / 1000) > maxSeconds ) {
-			System.out.println("### Requesting stop...");
+			if ( verbose ) System.out.println("### Requesting stop...");
 			ast.requestStop();
 		}
 	}

### 2008-05-05 ij3d/ImageCanvas3D.java
Fixed strange keyevent handling when toggling planes in orthoslice view.
Only process KeyPressed(), return immediately on keyReleased() and keyTyped().

@@ -24,8 +24,8 @@ import java.awt.Point;
 import java.awt.Color;
 import java.awt.Font;
 
-import java.util.Set;
-import java.util.HashSet;
+import java.util.Map;
+import java.util.HashMap;
 
 
 import ij.process.ByteProcessor;
@@ -34,16 +34,18 @@ import ij.gui.ImageCanvas;
 import ij.ImagePlus;
 import ij.gui.Roi;
 
-public class ImageCanvas3D extends Canvas3D {
+public class ImageCanvas3D extends Canvas3D implements KeyListener {
 
 	private RoiImagePlus roiImagePlus;
 	private ImageCanvas roiImageCanvas;
-	private Set<Integer> keysDown = new HashSet<Integer>();
+	private Map<Integer, Long> pressed, released; 
 
 	private class RoiImagePlus extends ImagePlus {
 		public RoiImagePlus(String title, ByteProcessor ip) {
 			super();
 			setProcessor(title, ip);
+			pressed = new HashMap<Integer, Long>();
+			released = new HashMap<Integer, Long>();
 		}
 
 		public ImageCanvas getCanvas() {
@@ -63,6 +65,7 @@ public class ImageCanvas3D extends Canvas3D {
 					super.mousePressed(e);
 			}
 		};
+		roiImageCanvas.removeKeyListener(ij.IJ.getInstance());
 		roiImageCanvas.disablePopupMenu(true);
 		
 		addListeners();
@@ -114,25 +117,11 @@ public class ImageCanvas3D extends Canvas3D {
 				render();
 			}
 		});
-		addKeyListener(new KeyListener() {
-			public void keyPressed(KeyEvent e) {
-				keysDown.add(e.getKeyCode());
-			}
-
-			public void keyReleased(KeyEvent e) {
-				keysDown.remove(e.getKeyCode());
-			}
-
-			public void keyTyped(KeyEvent e) {}
-		});
+		addKeyListener(this);
 	} 
 
 	public Roi getRoi() {
 		return roiImagePlus.getRoi();
-	}
-
-	public boolean isKeyDown(int keyCode) {
-		return keysDown.contains(keyCode);
 	}
 
 	public void render() {
@@ -141,46 +130,39 @@ public class ImageCanvas3D extends Canvas3D {
 		startRenderer();
 	}
 
-	private String status = "";
-	private Color color = Color.WHITE;
+	/*
+	 * Needed for the isKeyDown() method. Problem:
[…]

### 2008-05-05 ij3d/MouseBehavior.java

@@ -136,7 +136,7 @@ public class MouseBehavior extends Behavior {
 	public void doProcess(KeyEvent e) {
 		int id = e.getID();
 
-		if(id == KeyEvent.KEY_RELEASED)
+		if(id == KeyEvent.KEY_RELEASED || id == KeyEvent.KEY_TYPED)
 			return;
 
 		Content c = univ.getSelected();
@@ -151,34 +151,32 @@ public class MouseBehavior extends Behavior {
 			axis = Renderer.Z_AXIS;
 		if(e.isShiftDown()) {
 			switch(code) {
-				case KeyEvent.VK_RIGHT:translate(c, 5, 0);break;
-				case KeyEvent.VK_LEFT:translate(c, -5, 0);break;
-				case KeyEvent.VK_UP: translate(c, 0, -5);break;
-				case KeyEvent.VK_DOWN: translate(c, 0, 5);break;
+				case KeyEvent.VK_RIGHT:translate(c,5,0);return;
+				case KeyEvent.VK_LEFT:translate(c,-5,0);return;
+				case KeyEvent.VK_UP: translate(c,0,-5);return;
+				case KeyEvent.VK_DOWN: translate(c,0,5);return;
 			}
 		} else if(e.isAltDown()) {
 			switch(code) {
-				case KeyEvent.VK_UP: zoom(c, 1); break;
-				case KeyEvent.VK_DOWN: zoom(c, -1); break;
+				case KeyEvent.VK_UP: zoom(c, 1); return;
+				case KeyEvent.VK_DOWN: zoom(c, -1); return;
 			}
 		} else if(c != null && c.getType() == Content.ORTHO && axis != -1) {
 			OrthoGroup og = (OrthoGroup)c.getContent();
 			switch(code) {
 				case KeyEvent.VK_RIGHT:
-				case KeyEvent.VK_UP: og.increase(axis); break;
+				case KeyEvent.VK_UP: og.increase(axis); return;
 				case KeyEvent.VK_LEFT:
-				case KeyEvent.VK_DOWN: og.decrease(axis); break;
-				case KeyEvent.VK_SPACE:
-					boolean b = og.isVisible(axis);
-					og.setVisible(axis, !b);
-					break;
+				case KeyEvent.VK_DOWN:og.decrease(axis); return;
+				case KeyEvent.VK_SPACE: og.setVisible(axis,
[…]

### 2008-05-05 ij3d/Image3DMenubar.java
Fixed 2 possibilities for runtime NullPointerExceptions.

@@ -617,8 +617,8 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 			if(t != null) {
 				c.setTransform(new Transform3D(t));
 				univ.fireTransformationFinished();
+				record(SET_TRANSFORM, affine2string(t));
 			}
-			record(SET_TRANSFORM, affine2string(t));
 		}
 
 		if(e.getSource() == properties) {
@@ -657,8 +657,8 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 			if(t != null) {
 				c.applyTransform(new Transform3D(t));
 				univ.fireTransformationFinished();
+				record(APPLY_TRANSFORM, affine2string(t));
 			}
-			record(APPLY_TRANSFORM, affine2string(t));
 		}
 
 		if(e.getSource() == saveTransform) {

### 2008-05-05 ij3d/Image3DMenubar.java
Allow to export a transformed image from the 3D universe.

@@ -24,6 +24,7 @@ import java.util.Map;
 
 import vib.PointList;
 import vib.BenesNamedPoint;
+import vib.InterpolatedImage;
 import vib.FastMatrix;
 
 import orthoslice.OrthoGroup;
@@ -34,6 +35,7 @@ import isosurface.MeshExporter;
 import isosurface.MeshEditor;
 
 import javax.vecmath.Color3f;
+import javax.vecmath.Matrix4d;
 import javax.media.j3d.View;
 import javax.media.j3d.Transform3D;
 
@@ -66,6 +68,7 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 	private MenuItem resetTransform;
 	private MenuItem applyTransform;
 	private MenuItem saveTransform;
+	private MenuItem exportTransformed;
 	private MenuItem exportObj;
 	private MenuItem exportDXF;
 	private MenuItem smoothMesh;
@@ -302,6 +305,12 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 		saveTransform = new MenuItem("Save Transform");
 		saveTransform.addActionListener(this);
 		transform.add(saveTransform);
+
+		transform.addSeparator();
+
+		exportTransformed= new MenuItem("Export transformed image");
+		exportTransformed.addActionListener(this);
+		transform.add(exportTransformed);
 
 		return transform;
 	}
@@ -678,6 +687,15 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 				record(SAVE_TRANSFORM, affine2string(matrix));
 		}
 
+		if(e.getSource() == exportTransformed) {
+			Content c = univ.getSelected();
+			if(c == null) {
+				IJ.error("Selection required");
+				return;
+			}
+			exportTransformed(c);
+		}
+
 		if (e.getSource() == pl_load) {
 			Content c = univ.getSelected();
 			if(c == null) {
@@ -817,6 +835,35 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 		if(gd.wasCanceled())
 			return;
 		univ.setSize((int)gd.getNextNumber(), (int)gd.getNextNumber());
+	}
+
+	public void exportTransformed(Content c) {
+		ImagePlus orig = c.getImage();
+		Transform3D t1 = new Transform3D();
+		c.getLocalTranslate().getTransform(t1);
[…]

### 2008-05-05 ij3d/Image3DMenubar.java
Export transformed image in extra thread.

@@ -688,12 +688,16 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 		}
 
 		if(e.getSource() == exportTransformed) {
-			Content c = univ.getSelected();
+			final Content c = univ.getSelected();
 			if(c == null) {
 				IJ.error("Selection required");
 				return;
 			}
-			exportTransformed(c);
+			new Thread(new Runnable() {
+				public void run() {
+					exportTransformed(c);
+				}
+			}).start();
 		}
 
 		if (e.getSource() == pl_load) {
@@ -839,6 +843,11 @@ public class Image3DMenubar extends MenuBar implements ActionListener,
 
 	public void exportTransformed(Content c) {
 		ImagePlus orig = c.getImage();
+		if(orig == null) {
+			IJ.error("No greyscale image exists for "
+				+ c.getName());
+			return;
+		}
 		Transform3D t1 = new Transform3D();
 		c.getLocalTranslate().getTransform(t1);
 		Transform3D t2 = new Transform3D();

### 2008-05-05 tracing/AutoPoint.java
Changes mostly to improve memory handling; also write wavefront .obj files

@@ -22,16 +22,13 @@
 package tracing;
 
 public class AutoPoint {
-	public int x;
-	public int y;
-	public int z;
-	public boolean overThreshold = false;
-	public AutoPoint [] predecessors;
+	public short x;
+	public short y;
+	public short z;
 	public AutoPoint(int x,int y,int z) {
-		this.x = x;
-		this.y = y;
-		this.z = z;
-		this.predecessors = null;
+		this.x = (short)x;
+		this.y = (short)y;
+		this.z = (short)z;
 	}
 	@Override
 	public String toString() {
@@ -43,32 +40,5 @@ public class AutoPoint {
 		// System.out.println("Testing equality between "+this+" and "+op);
 		boolean result = (this.x == op.x) && (this.y == op.y) && (this.z == op.z);
 		return result;
-	}
-	public void addPredecessor(AutoPoint p) {
-		if( predecessors == null ) {
-			predecessors = new AutoPoint[1];
-			predecessors[0] = p;
-		} else {
-			for( int i = 0; i < predecessors.length; ++i )
-				if( p.equals( predecessors[i] ) )
-					return;
-			AutoPoint [] n = new AutoPoint[predecessors.length+1];
-			System.arraycopy(predecessors,0,n,0,predecessors.length);
-			n[predecessors.length] = p;
-			predecessors = n;
-		}
-	}
-	public void addPredecessors(AutoPoint [] newPredecessors) {
-		if( newPredecessors == null )
-			return;
-		if( predecessors == null ) {
-			predecessors = new AutoPoint[newPredecessors.length];
-			System.arraycopy(newPredecessors,0,predecessors,0,newPredecessors.length);
-		} else {
-			AutoPoint [] n = new AutoPoint[predecessors.length+newPredecessors.length];
-			System.arraycopy(predecessors,0,n,0,predecessors.length);
-			System.arraycopy(newPredecessors,0,n,predecessors.length,newPredecessors.length);
-			predecessors = n;
-		}
 	}
 }

### 2008-05-05 tracing/AutoSearchThread.java

@@ -91,11 +91,9 @@ public class AutoSearchThread extends SearchThread {
 	protected void addingNode( SearchNode n ) {
 		if( tubeValues[n.z][n.y*width+n.x] > tubenessThreshold ) {
 			AutoPoint p=new AutoPoint(n.x,n.y,n.z);
-			p.overThreshold = true;
 			destinations.add(p);
 		} else if( null != previousPathGraph.get(n.x,n.y,n.z) ) {
 			AutoPoint p=new AutoPoint(n.x,n.y,n.z);
-			p.overThreshold = false;
 			destinations.add(p);
 		}
 	}

### 2008-05-05 tracing/Auto_Tracer.java

@@ -28,7 +28,8 @@ import ij.ImageStack;
 import ij.io.FileInfo;
 import ij.io.FileSaver;
 import ij.plugin.PlugIn;
-import java.io.File;
+import ij.measure.Calibration;
+import java.io.*;
 import java.util.PriorityQueue;
 import java.util.Comparator;
 import java.util.ArrayList;
@@ -141,22 +142,29 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			}
 		}
 
-		SinglePathsGraph completePaths = new SinglePathsGraph(width,height,depth);
+		Calibration c = image.getCalibration();
+		SinglePathsGraph completePaths = new SinglePathsGraph(width,height,depth,c.pixelWidth,c.pixelHeight,c.pixelDepth);
 
-		int maxLoops = 4;
+		int maxLoops = 400;
 		int loopsDone = 0;
 
 		while( mostTubelikePoints.size() > 0 ) {
 
-			if( maxLoops >= 0 && loopsDone > maxLoops )
+			if( maxLoops >= 0 && loopsDone >= maxLoops )
 				break;
-			
+
 			System.out.println("=== Priority queue now has: "+mostTubelikePoints.size());
 
 			// Now get the most tubelike point:
 			AutoPoint startPoint=mostTubelikePoints.poll();
 
 			System.out.println("Got point "+startPoint);
+
+			if( loopsDone == 0 ) {
+				startPoint.x = 221;
+				startPoint.y = 155;
+				startPoint.z = 37;
+			}
 			
 			ast = new AutoSearchThread( image, /* original image */
 						    tubeValues, /* the "tubeness" filtered image */
@@ -179,6 +187,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				ast.join();
 			} catch( InterruptedException e ) { }
 
+			canvas.removeSearchThread(ast);
+
 			// Now start the pruning:
 
 			ArrayList<AutoPoint> destinations = ast.getDestinations();
@@ -200,37 +210,44 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 				int lastIndex = path.size() - 1;
 				
-				for( int i = 0; i < path.size(); ++i  ) {
+				if( minimumPointsOnPath >= 0 && path.size() < minimumPointsOnPath ) {
 
-					if ( verbose ) System.out.print(".");
-					if ( verbose ) System.out.flush();
[…]

### 2008-05-05 tracing/SinglePathsGraph.java

@@ -22,17 +22,28 @@
 package tracing;
 
 import java.util.Hashtable;
+import java.util.Set;
+import java.util.Iterator;
+import java.io.*;
+import java.util.zip.GZIPOutputStream;
 
 public class SinglePathsGraph {
 	
 	int width, height, depth;
-	
-	public SinglePathsGraph( int width, int height, int depth ) {
+	float spacing_x, spacing_y, spacing_z;
+
+	public SinglePathsGraph( int width, int height, int depth, double spacing_x, double spacing_y, double spacing_z ) {
 		this.width = width;
 		this.height = height;
 		this.depth = depth;
+		this.spacing_x = (float)spacing_x;
+		this.spacing_y = (float)spacing_y;
+		this.spacing_z = (float)spacing_z;
+		this.directedLinksCapacity = 256;
+		this.directedLinksInArray = 0;
+		this.directedLinks = new int[directedLinksCapacity*2];
 	}
-	
+
 	// For fast lookup from positions:
 	Hashtable<Integer,AutoPoint> fromPosition=new Hashtable<Integer,AutoPoint>();
 	
@@ -41,15 +52,149 @@ public class SinglePathsGraph {
 		return fromPosition.get(k);
 	}
 	
-	public void addPoint( AutoPoint p ) {
-		int k = p.x + p.y * width + p.z * width * height;
+	public void addPoint( AutoPoint p, AutoPoint previous ) {
+		int k = (int)p.x + (int)p.y * width + (int)p.z * width * height;
 		AutoPoint existingPoint=fromPosition.get( k );
 		if( existingPoint == null ) {
 			fromPosition.put(k,p);
-		} else {
-			// "merge" this point with the exisiting one -
-			// i.e. just add the predecessors:
-			existingPoint.addPredecessors(p.predecessors);
 		}
-	}	
-}
+		if( previous != null ) {
+			int k_previous = (int)previous.x + (int)previous.y * width + (int)previous.z * width * height;
+			addLink( k_previous, k );
+		}
+	}
+
+	// This all turns out to be very memory hungry, so keep track
+	// of the links between points here:
+
+	int directedLinksInArray;
+	int directedLinksCapacity;
+
+	int [] directedLinks;
+
+	void addLink( int from_k, int to_k ) {
+		if( directedLinksInArray == directedLinksCapacity ) {
+			int newCapacity = (directedLinksCapacity * 5) / 4;
+			int [] n = new int[newCapacity*2];
[…]

### 2008-05-05 vib/InterpolatedImage.java
Add Dirichlet Edge plugin
This is the plugin demonstrating an information theoretic approach to
edge detection (to be published).

Signed-off-by: Johannes Schindelin <johannes.schindelin@gmx.de>

@@ -52,6 +52,10 @@ public class InterpolatedImage {
 		}
 	}
 
+	public int getWidth() { return w; }
+	public int getHeight() { return h; }
+	public int getDepth() { return d; }
+
 	protected InterpolatedImage() {
 	}
 

### 2008-05-05 tracing/Auto_Tracer.java
Make the defaults slightly more sensible.

@@ -145,7 +145,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		Calibration c = image.getCalibration();
 		SinglePathsGraph completePaths = new SinglePathsGraph(width,height,depth,c.pixelWidth,c.pixelHeight,c.pixelDepth);
 
-		int maxLoops = 400;
+		int maxLoops = 100;
 		int loopsDone = 0;
 
 		while( mostTubelikePoints.size() > 0 ) {
@@ -319,13 +319,14 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 	long timeStarted;
 
-	int maxNodes = 100000;
-	// int maxNodes = 4000;
-	int maxSeconds = 10;
-	float tubenessThreshold = 41.0f;
-	float minimumRollingMean = 10.0f;
+	int maxNodes = 22000; // Takes about 10 seconds to do this on a 1.8GHz Duron
+	int maxSeconds = 120;
+	float tubenessThreshold = 18f;
+	float minimumRollingMean = 5.0f;
 	int rollingLength = 4;
 
+	// If you only want to add paths that have more than a certain
+	// number of points in them, change this...
 	int minimumPointsOnPath = -1;
 
 	/* This is a (hopefully accurate) description of the algorithm
@@ -450,7 +451,10 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 	 * e.g. TracerThreed.getResult() */
 	
 	public void finished( SearchThread source, boolean success ) {
-		
+		long currentTime = System.currentTimeMillis();
+		long secondsSinceStarted = (currentTime - timeStarted) / 1000;
+		// Just log how many nodes were explored in that time:
+		System.out.println("  "+(source.open_from_start.size()+source.closed_from_start.size())+" nodes in "+secondsSinceStarted+" seconds");
 	}
 	
 	/* This reports the current status of the thread, which may be:

### 2008-05-05 tracing/Auto_Tracer.java
Load parameters from .thresholds files or put up a generic dialog

@@ -4,17 +4,17 @@
 
 /*
   This file is part of the ImageJ plugin "Auto Tracer".
-  
+
   The ImageJ plugin "Auto Tracer" is free software; you can
   redistribute it and/or modify it under the terms of the GNU General
   Public License as published by the Free Software Foundation; either
   version 3 of the License, or (at your option) any later version.
-  
+
   The ImageJ plugin "Auto Tracer" is distributed in the hope that it
   will be useful, but WITHOUT ANY WARRANTY; without even the implied
   warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
   See the GNU General Public License for more details.
-  
+
   You should have received a copy of the GNU General Public License
   along with this program.  If not, see <http://www.gnu.org/licenses/>.
 */
@@ -29,6 +29,7 @@ import ij.io.FileInfo;
 import ij.io.FileSaver;
 import ij.plugin.PlugIn;
 import ij.measure.Calibration;
+import ij.gui.GenericDialog;
 import java.io.*;
 import java.util.PriorityQueue;
 import java.util.Comparator;
@@ -68,6 +69,50 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
                 return new AutoTracerCanvas( imagePlus, this, plane, null );
         }
 
+	static class AutoTracerParameters {
+		float tubenessThreshold;
+		float minimumRollingMean;
+	}
+
+	AutoTracerParameters loadParameters(String filename) {
+		float tubenessThreshold = Float.MIN_VALUE;
+		float minimumRollingMean = Float.MIN_VALUE;
+		try {
+			BufferedReader br = new BufferedReader(new FileReader(filename));
+			String lastLine;
+			while( null != (lastLine = br.readLine()) ) {
+				lastLine = lastLine.trim();
+				if( lastLine.length() == 0 )
+					continue;
+				if( lastLine.length() < 4 )
+					return null;
+				if( lastLine.startsWith("t: ") ) {
+					String rest=lastLine.substring(3);
+					System.out.println("   For tubenessThreshold got string: '"+rest+"'");
+					try {
+						tubenessThreshold = Float.parseFloat(rest);
+					} catch( NumberFormatException e ) {
+						return null;
+					}
+				} else if( lastLine.startsWith("m: ") ) {
+					String rest=lastLine.substring(3);
[…]

### 2008-05-06 tracing/Auto_Tracer.java
Output a more sensible filename

@@ -143,6 +143,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		String beforeExtension=originalFileName.substring(0, lastDot);
 		String tubesFileName=beforeExtension+".tubes.tif";
 		String thresholdsFileName=beforeExtension+".thresholds";
+		String outputFileName=beforeExtension+".traces.obj";
 		ImagePlus tubenessImage = null;
 		File tubesFile=new File(originalFileInfo.directory,tubesFileName);
 		if( tubesFile.exists() ) {
@@ -388,12 +389,12 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			++loopsDone;
 		}
 
-		String outputFilename = "/home/mark/test.obj";
+		File outputFile=new File(originalFileInfo.directory,outputFileName);
 
 		try {
-			completePaths.writeWavefrontObj(outputFilename);
+			completePaths.writeWavefrontObj(outputFile.getAbsolutePath());
 		} catch( IOException e ) {
-			IJ.error("Writing the Wavefront OBJ file '"+outputFilename+"' failed");
+			IJ.error("Writing the Wavefront OBJ file '"+outputFile.getAbsolutePath()+"' failed");
 			return;
 		}
 

### 2008-05-06 tracing/Auto_Tracer.java
remove() on the priority queue is O(n) so instead record the ones to skip in a HashSet

@@ -35,6 +35,7 @@ import java.util.PriorityQueue;
 import java.util.Comparator;
 import java.util.ArrayList;
 import java.util.Iterator;
+import java.util.HashSet;
 import util.BatchOpener;
 
 import java.awt.Color;
@@ -48,6 +49,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 	int height;
 	int depth;
 
+	HashSet<AutoPoint> skip;		
 	PriorityQueue<AutoPoint> mostTubelikePoints;
 	float [][] tubeValues;
 
@@ -243,7 +245,16 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			// Now get the most tubelike point:
 			AutoPoint startPoint=mostTubelikePoints.poll();
 
-			System.out.println("Got point "+startPoint);
+			if( skip.contains(startPoint) ) {
+				skip.remove(startPoint);
+				continue;
+			}
+			
+			System.out.println("=== Skip size is: "+skip.size());
+			System.out.println("=== Priority queue now has: "+mostTubelikePoints.size());
+			System.out.println("=== Loops done: "+loopsDone);
+			
+			System.out.println("  Got point "+startPoint+" with tubeness: "+tubeValues[startPoint.z][startPoint.y*width+startPoint.x]);
 
 			if( loopsDone == 0 ) {
 				startPoint.x = 221;
@@ -373,11 +384,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 					if ( verbose ) System.out.flush();
 
-					if( mostTubelikePoints.remove( toRemove ) ) {
-						if ( verbose ) System.out.print("-");
-					} else {
-						if ( verbose ) System.out.print("x");
-					}
+					skip.add( toRemove );
 				}
 
 				if (verbose) System.out.println("");

### 2008-05-06 tracing/Auto_Tracer.java
For some reason Java on my Mac doesn't like this @Override

@@ -127,7 +127,6 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			this.tubeValues = tubeValues;
 		}
 
-		@Override
 		public int compare( AutoPoint a, AutoPoint b ) {
 			return -Float.compare(tubeValues[a.z][a.y*width+a.x],tubeValues[b.z][b.y*width+b.x]);
 		}

### 2008-05-06 tracing/AutoSearchThread.java
Tiny cosmetic change

@@ -35,7 +35,7 @@ public class AutoSearchThread extends SearchThread {
 
 	SinglePathsGraph previousPathGraph;
 
-	ArrayList<AutoPoint> destinations = new ArrayList(512);
+	ArrayList<AutoPoint> destinations = new ArrayList<AutoPoint>(512);
 
 	public ArrayList<AutoPoint> getDestinations() {
 		return destinations;

### 2008-05-06 tracing/Auto_Tracer.java
Recreate the priority queue (not adding those in skip) if memory is low

@@ -133,6 +133,41 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 	}
 
+	public void recreatePriorityQueue( boolean checkSkip ) {
+	
+		System.out.println("  [Recreating Priority Queue]");
+		mostTubelikePoints=new PriorityQueue<AutoPoint>(512,new TubenessComparator(width,height,depth,tubeValues));
+		System.gc();
+		
+		for(int z=0;z<depth;++z) {
+			for(int y=0;y<height;++y) {
+				for(int x=0;x<width;++x) {
+					if( tubeValues[z][y*width+x] > tubenessThreshold ) {
+						AutoPoint p=new AutoPoint(x,y,z);
+						if (checkSkip) {
+							if( skip.contains(p) )
+								skip.remove(p);
+							else
+								mostTubelikePoints.add(p);
+						} else
+							mostTubelikePoints.add(p);
+					}
+				}
+			}
+		}
+
+		if( checkSkip ) {
+			if( skip.size() != 0 ) {
+				throw new RuntimeException("The skip HashSet was not empty after recreating (BUG) - had "+skip.size()+" elements");
+			}
+			skip = null;
+			skip = new HashSet<AutoPoint>();
+			System.gc();
+		}
+		System.out.println("  [Done]");
+	}
+
+	
 	boolean verbose = false;
 
 	public void autoTrace( ImagePlus image ) {
@@ -215,18 +250,14 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			tubeValues[z]=(float[])tubeStack.getPixels(z+1);
 		}
 
-		mostTubelikePoints=new PriorityQueue<AutoPoint>(512,new TubenessComparator(width,height,depth,tubeValues));
+		skip = new HashSet<AutoPoint>();
 
-		// Add all of those points to the priority queue....
-		for(int z=0;z<depth;++z) {
-			for(int y=0;y<height;++y) {
-				for(int x=0;x<width;++x) {
-					if( tubeValues[z][y*width+x] > tubenessThreshold ) {
-						mostTubelikePoints.add(new AutoPoint(x,y,z));
-					}
-				}
-			}
-		}
+		recreatePriorityQueue(false);
+		
+		System.out.println("Initial points: "+mostTubelikePoints.size());
+
+		if( false )
+			return;
 
 		Calibration c = image.getCalibration();
[…]

### 2008-05-06 tracing/AutoPoint.java
Add a plausible hashCode for AutoPoint

@@ -41,4 +41,8 @@ public class AutoPoint {
 		boolean result = (this.x == op.x) && (this.y == op.y) && (this.z == op.z);
 		return result;
 	}
+	@Override
+	public int hashCode() {
+		return (int)x + (int)y * (1 << 11) + (int)z * (1 << 22);
+	}
 }

### 2008-05-06 tracing/Auto_Tracer.java
Temporarily make it an error to not have a parameters file

@@ -217,6 +217,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			tubenessThreshold = p.tubenessThreshold;
 			minimumRollingMean = p.minimumRollingMean;
 		} else {
+			if (true)
+				throw new RuntimeException("Tried to create a generic dialog!");
 			// Pop up a GenericDialog to ask:
 			GenericDialog gd = new GenericDialog("Auto Tracer");
 			gd.addNumericField( "Tubeness threshold for destinations", tubenessThreshold, 2 );

### 2008-05-06 tracing/Auto_Tracer.java
Very slight speed improvement by making destinationsToPrune a HashSet

@@ -388,7 +388,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				AutoPoint current = null;
 				AutoPoint last = null;
 
-				ArrayList<AutoPoint> destinationsToPrune = new ArrayList<AutoPoint>();
+				HashSet<AutoPoint> destinationsToPrune = new HashSet<AutoPoint>();
 
 				for( int i = 0; i <= lastIndex; ++i ) {
 

### 2008-05-06 tracing/Auto_Tracer.java
Turn off the loop limit

@@ -264,7 +264,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		Calibration c = image.getCalibration();
 		SinglePathsGraph completePaths = new SinglePathsGraph(width,height,depth,c.pixelWidth,c.pixelHeight,c.pixelDepth);
 
-		int maxLoops = 100;
+		int maxLoops = -1;
 		int loopsDone = 0;
 
 		while( mostTubelikePoints.size() > 0 ) {

### 2008-05-06 tracing/Auto_Tracer.java
Extra verbosity

@@ -327,6 +327,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			// Now start the pruning:
 
 			ArrayList<AutoPoint> destinations = ast.getDestinations();
+			System.out.println("  === Destinations: "+destinations.size());
 			if ( verbose ) System.out.print("  === Destinations: "+destinations.size()+" ");
 			if ( verbose ) System.out.flush();
 

### 2008-05-06 tracing/Auto_Tracer.java
The priority queue might be empty after recreating

@@ -278,8 +278,11 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			
 			// This should help the memory usage somewhat:
 			System.out.println("=== Memory usage: "+percentUsed+"%");
-			if( percentUsed > 95 )
+			if( percentUsed > 95 ) {
 				recreatePriorityQueue(true);
+				if( mostTubelikePoints.size() == 0 )
+					break;
+			}
 
 			// Now get the most tubelike point:
 			AutoPoint startPoint=mostTubelikePoints.poll();

### 2008-05-06 tracing/Auto_Tracer.java
Took out some debugging code accidentally left in

@@ -298,12 +298,6 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			
 			System.out.println("  Got point "+startPoint+" with tubeness: "+tubeValues[startPoint.z][startPoint.y*width+startPoint.x]);
 
-			if( loopsDone == 0 ) {
-				startPoint.x = 221;
-				startPoint.y = 155;
-				startPoint.z = 37;
-			}
-
 			ast = new AutoSearchThread( image, /* original image */
 						    tubeValues, /* the "tubeness" filtered image */
 						    startPoint, /* the point to start the search from */

### 2008-05-06 tracing/Auto_Tracer.java
Make the open and closed nodes a different colour

@@ -306,7 +306,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 			timeStarted = System.currentTimeMillis();
 
-			ast.setDrawingColors( Color.CYAN, Color.CYAN );
+			ast.setDrawingColors( Color.BLUE, Color.CYAN );
 			ast.setDrawingThreshold( -1 );
 
 			ast.addProgressListener(this);

### 2008-05-06 tracing/Auto_Tracer.java
Move the memory check to the end of the file

@@ -272,18 +272,6 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			if( maxLoops >= 0 && loopsDone >= maxLoops )
 				break;
 
-                	long freeMem = Runtime.getRuntime().freeMemory();
-		        long totMem = Runtime.getRuntime().totalMemory();
-			int percentUsed = (int)(((totMem-freeMem)*100) / totMem);
-			
-			// This should help the memory usage somewhat:
-			System.out.println("=== Memory usage: "+percentUsed+"%");
-			if( percentUsed > 95 ) {
-				recreatePriorityQueue(true);
-				if( mostTubelikePoints.size() == 0 )
-					break;
-			}
-
 			// Now get the most tubelike point:
 			AutoPoint startPoint=mostTubelikePoints.poll();
 
@@ -428,7 +416,22 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			}
 
 			ast = null;
+
 			System.gc();
+
+			/* Check the memory usage, and recreate the
+			 * PriorityQueue if it's going too high: */
+
+                	long freeMem = Runtime.getRuntime().freeMemory();
+		        long totMem = Runtime.getRuntime().totalMemory();
+			int percentUsed = (int)(((totMem-freeMem)*100) / totMem);
+			
+			System.out.println("=== Memory usage: "+percentUsed+"%");
+			if( percentUsed > 95 ) {
+				recreatePriorityQueue(true);
+				if( mostTubelikePoints.size() == 0 )
+					break;
+			}
 
 			++loopsDone;
 		}

### 2008-05-06 tracing/SinglePathsGraph.java
Write out image metadata in the comments of the .obj file

@@ -95,6 +95,16 @@ public class SinglePathsGraph {
 
 		PrintWriter pw = new PrintWriter(new OutputStreamWriter(new FileOutputStream(outputFilename),"UTF-8"));
 
+		// It's a bit ugly encoding this data in the comments,
+		// but it's a useful enough stop gap measure:
+		pw.println("# width: "+width);
+		pw.println("# height: "+height);
+		pw.println("# depth: "+depth);
+		pw.println("# spacing_x: "+spacing_x);
+		pw.println("# spacing_y: "+spacing_y);
+		pw.println("# spacing_z: "+spacing_z);
+		
+
 		/* First build a map from keys to wavefront indices: */
 
 		Hashtable<Integer,Integer> wavefrontToKey=new Hashtable<Integer,Integer>();
@@ -123,11 +133,6 @@ public class SinglePathsGraph {
 		pw.print("g");
 
 		pw.close();
-
-		
-		
-
-
 	}
 
 

### 2008-05-06 tracing/Auto_Tracer.java
Move to the slice at the center of each new search

@@ -286,6 +286,9 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			
 			System.out.println("  Got point "+startPoint+" with tubeness: "+tubeValues[startPoint.z][startPoint.y*width+startPoint.x]);
 
+			// Move to that slice, just for presentation purposes:
+			image.setSlice(startPoint.z+1);
+
 			ast = new AutoSearchThread( image, /* original image */
 						    tubeValues, /* the "tubeness" filtered image */
 						    startPoint, /* the point to start the search from */

### 2008-05-06 tracing/Auto_Tracer.java
Comment fix, and also recreate the queue every 50 loops

@@ -403,7 +403,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 					last = current;
 				}
 
-				// Now remove all the points genuinely found in this search:
+				// Now remove all the destinations
+				// genuinely found in this search:
 
 				for( Iterator<AutoPoint> itRemove = destinationsToPrune.iterator();
 				     itRemove.hasNext(); ) {
@@ -423,14 +424,17 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			System.gc();
 
 			/* Check the memory usage, and recreate the
-			 * PriorityQueue if it's going too high: */
+			 * PriorityQueue if it's going too high (and
+			 * every 50 loops in any case).  It's not
+			 * clear how trustworthy these memory
+			 * statistics are, unfortunately. */
 
                 	long freeMem = Runtime.getRuntime().freeMemory();
 		        long totMem = Runtime.getRuntime().totalMemory();
 			int percentUsed = (int)(((totMem-freeMem)*100) / totMem);
 			
 			System.out.println("=== Memory usage: "+percentUsed+"%");
-			if( percentUsed > 95 ) {
+			if( (percentUsed > 95) || ((loopsDone % 50) == 49) ) {
 				recreatePriorityQueue(true);
 				if( mostTubelikePoints.size() == 0 )
 					break;

### 2008-05-06 tracing/Auto_Tracer.java
Rename "skip" to "done" and fix a bug that would make this never terminate

@@ -49,7 +49,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 	int height;
 	int depth;
 
-	HashSet<AutoPoint> skip;		
+	HashSet<AutoPoint> done;		
 	PriorityQueue<AutoPoint> mostTubelikePoints;
 	float [][] tubeValues;
 
@@ -133,7 +133,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 	}
 
-	public void recreatePriorityQueue( boolean checkSkip ) {
+	public void recreatePriorityQueue( boolean checkDone ) {
 	
 		System.out.println("  [Recreating Priority Queue]");
 		mostTubelikePoints=new PriorityQueue<AutoPoint>(512,new TubenessComparator(width,height,depth,tubeValues));
@@ -144,25 +144,14 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				for(int x=0;x<width;++x) {
 					if( tubeValues[z][y*width+x] > tubenessThreshold ) {
 						AutoPoint p=new AutoPoint(x,y,z);
-						if (checkSkip) {
-							if( skip.contains(p) )
-								skip.remove(p);
-							else
+						if (checkDone) {
+							if( ! done.contains(p) )
 								mostTubelikePoints.add(p);
 						} else
 							mostTubelikePoints.add(p);
 					}
 				}
 			}
-		}
-
-		if( checkSkip ) {
-			if( skip.size() != 0 ) {
-				throw new RuntimeException("The skip HashSet was not empty after recreating (BUG) - had "+skip.size()+" elements");
-			}
-			skip = null;
-			skip = new HashSet<AutoPoint>();
-			System.gc();
 		}
 		System.out.println("  [Done]");
 	}
@@ -252,7 +241,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			tubeValues[z]=(float[])tubeStack.getPixels(z+1);
 		}
 
-		skip = new HashSet<AutoPoint>();
+		done = new HashSet<AutoPoint>();
 
 		recreatePriorityQueue(false);
 		
@@ -275,12 +264,11 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			// Now get the most tubelike point:
 			AutoPoint startPoint=mostTubelikePoints.poll();
 
-			if( skip.contains(startPoint) ) {
[…]

### 2008-05-06 tracing/Load_Auto_Traces.java
Support for loading auto traces files

@@ -1,0 +1,177 @@
+/* -*- mode: java; c-basic-offset: 8; indent-tabs-mode: t; tab-width: 8 -*- */
+
+/* Copyright 2006, 2007 Mark Longair */
+
+/*
+  This file is part of the ImageJ plugin "Auto Tracer".
+
+  The ImageJ plugin "Auto Tracer" is free software; you can
+  redistribute it and/or modify it under the terms of the GNU General
+  Public License as published by the Free Software Foundation; either
+  version 3 of the License, or (at your option) any later version.
+
+  The ImageJ plugin "Auto Tracer" is distributed in the hope that it
+  will be useful, but WITHOUT ANY WARRANTY; without even the implied
+  warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+  See the GNU General Public License for more details.
+
+  You should have received a copy of the GNU General Public License
+  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+package tracing;
+
+import ij.*;
+import ij.plugin.*;
+import ij.gui.*;
+import ij.io.*;
+import ij.process.*;
+
+import java.io.*;
+import java.awt.*;
+import java.util.StringTokenizer;
+
+public class Load_Auto_Traces implements PlugIn {
+
+	public int getNonNegativeIntAfterPrefix( String line, String prefix ) {
+		try {
+			String s=line.substring(prefix.length());
+			return Integer.parseInt(s);
+		} catch( NumberFormatException e ) {
+			return -1;
+		}
+	}
+
+	public float getNonNegativeFloatAfterPrefix( String line, String prefix ) {
+		try {
+			String s=line.substring(prefix.length());
+			return Float.parseFloat(s);
+		} catch( NumberFormatException e ) {
+			return -1;
+		}
+	}
+
+	public void run( String ignored ) {
+
+                OpenDialog od;
+		
+                od = new OpenDialog("Select traces.obj file...",
+                                    null,
+                                    null );
+		
+                String fileName = od.getFileName();
+                String directory = od.getDirectory();
+		
+                if( fileName == null )
+			return;
+		
+		System.out.println("Got "+fileName);
+
+		int width = -1, height = -1, depth = -1;
[…]

### 2008-05-06 tracing/Auto_Tracer.java
Add a macro option to turn off live display of progress...

@@ -25,6 +25,7 @@ import features.Tubeness_;
 import ij.IJ;
 import ij.ImagePlus;
 import ij.ImageStack;
+import ij.Macro;
 import ij.io.FileInfo;
 import ij.io.FileSaver;
 import ij.plugin.PlugIn;
@@ -275,7 +276,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			System.out.println("  Got point "+startPoint+" with tubeness: "+tubeValues[startPoint.z][startPoint.y*width+startPoint.x]);
 
 			// Move to that slice, just for presentation purposes:
-			image.setSlice(startPoint.z+1);
+			if( liveDisplay )
+				image.setSlice(startPoint.z+1);
 
 			ast = new AutoSearchThread( image, /* original image */
 						    tubeValues, /* the "tubeness" filtered image */
@@ -290,7 +292,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 			ast.addProgressListener(this);
 
-			canvas.addSearchThread(ast);
+			if( liveDisplay )
+				canvas.addSearchThread(ast);
 
 			ast.start();
 
@@ -298,7 +301,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 				ast.join();
 			} catch( InterruptedException e ) { }
 
-			canvas.removeSearchThread(ast);
+			if( liveDisplay )
+				canvas.removeSearchThread(ast);
 
 			// Now start the pruning:
 
@@ -528,6 +532,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 	 */
 
+	boolean liveDisplay = true;
+
 	public void run(String arg0) {
 
 		ImagePlus image = IJ.getImage();
@@ -544,13 +550,22 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		if( pointsInImage >= Integer.MAX_VALUE ) {
 			IJ.error("This plugin currently only works with images with less that "+Integer.MAX_VALUE+" points.");
 			return;
-		 }
+		}
 
+		String macroOptions = Macro.getOptions();
+		if( macroOptions != null ) {
+			String liveValue=Macro.getValue(macroOptions,"live","");
+			String lower = liveValue.toLowerCase();
[…]

### 2008-05-06 tracing/Auto_Tracer.java
Added an overall time limit (of 15 minutes)

@@ -230,6 +230,8 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			}
 		}
 
+		totalTimeStarted = System.currentTimeMillis();
+
 		System.out.println("Now using tubenessThreshold: "+tubenessThreshold);
 		System.out.println("     and minimumRollingMean: "+minimumRollingMean);
 
@@ -259,6 +261,11 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 
 		while( mostTubelikePoints.size() > 0 ) {
 
+			long currentTime = System.currentTimeMillis();
+			long secondsSinceStart = (currentTime - totalTimeStarted) / 1000;
+			if( secondsSinceStart > totalTimeLimitSeconds )
+				break;
+
 			if( maxLoops >= 0 && loopsDone >= maxLoops )
 				break;
 
@@ -285,7 +292,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 						    tubenessThreshold,
 						    completePaths );
 
-			timeStarted = System.currentTimeMillis();
+			threadTimeStarted = currentTime;
 
 			ast.setDrawingColors( Color.BLUE, Color.CYAN );
 			ast.setDrawingThreshold( -1 );
@@ -443,12 +450,14 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 			IJ.error("Writing the Wavefront OBJ file '"+outputFile.getAbsolutePath()+"' failed");
 			return;
 		}
-
 	}
 
 	AutoSearchThread ast;
 
-	long timeStarted;
+	long totalTimeLimitSeconds = 1 * 60;
+	long totalTimeStarted;
+
+	long threadTimeStarted;
 
 	int maxNodes = 22000; // Takes about 10 seconds to do this on a 1.8GHz Duron
 	int maxSeconds = 120;
@@ -582,7 +591,7 @@ public class Auto_Tracer extends ThreePanes implements PlugIn, PaneOwner, Search
 		// Also check whether we're over the requested number
 		// of iterations or time:
 		long currentTime = System.currentTimeMillis();
-		long timeSinceStarted = currentTime - timeStarted;
+		long timeSinceStarted = currentTime - threadTimeStarted;
 		if( (inOpen + inClosed) > maxNodes || (timeSinceStarted / 1000) > maxSeconds ) {
[…]

### 2008-05-06 tracing/SinglePathsGraph.java
Change to using a HashSet to track directed links

@@ -39,9 +39,6 @@ public class SinglePathsGraph {
 		this.spacing_x = (float)spacing_x;
 		this.spacing_y = (float)spacing_y;
 		this.spacing_z = (float)spacing_z;
-		this.directedLinksCapacity = 256;
-		this.directedLinksInArray = 0;
-		this.directedLinks = new int[directedLinksCapacity*2];
 	}
 
 	// For fast lookup from positions:
@@ -64,25 +61,16 @@ public class SinglePathsGraph {
 		}
 	}
 
-	// This all turns out to be very memory hungry, so keep track
-	// of the links between points here:
+	/* This all turns out to be very memory hungry, so keep track
+	   of the links between points here, encoded into a long for
+	   each link:
+	 */
 
-	int directedLinksInArray;
-	int directedLinksCapacity;
-
-	int [] directedLinks;
+	HashSet<Long> directedLinks = new HashSet<Long>();
 
 	void addLink( int from_k, int to_k ) {
-		if( directedLinksInArray == directedLinksCapacity ) {
-			int newCapacity = (directedLinksCapacity * 5) / 4;
-			int [] n = new int[newCapacity*2];
-			System.arraycopy(directedLinks,0,n,0,directedLinksInArray*2);
-			directedLinksCapacity = newCapacity;
-			directedLinks = n;
-		}
-		directedLinks[ directedLinksInArray * 2 ] = from_k;
-		directedLinks[ directedLinksInArray * 2 + 1 ] = to_k;
-		++ directedLinksInArray;
+		long toAdd = (from_k << 32) + to_k;
+		directedLinks.add(toAdd);
 	}
 
 	void addLink( int from_x, int from_y, int from_z, int to_x, int to_y, int to_z ) {
@@ -122,12 +110,14 @@ public class SinglePathsGraph {
 			++counter;
 		}
 		
-		for( int i = 0; i < directedLinksInArray; ++i ) {
-			int from_k = directedLinks[i*2];
-			int to_k = directedLinks[i*2+1];
+		for( Iterator<Long> links = directedLinks.iterator();
+		     links.hasNext(); ) {
+			long l = links.next();
+			long from = l >> 32;
+			long to = l & 0xFFFFFFFF;
 			int from_vertex = wavefrontToKey.get(from_k);
 			int to_vertex = wavefrontToKey.get(to_k);
-			pw.println("l "+from_vertex+" "+to_vertex);
+			pw.println("l "+from_vertex+" "+to_vertex);			
 		}
 
 		pw.print("g");

### 2008-05-06 tracing/SinglePathsGraph.java
Fix for broken compilation

@@ -22,6 +22,7 @@
 package tracing;
 
 import java.util.Hashtable;
+import java.util.HashSet;
 import java.util.Set;
 import java.util.Iterator;
 import java.io.*;
@@ -115,6 +116,8 @@ public class SinglePathsGraph {
 			long l = links.next();
 			long from = l >> 32;
 			long to = l & 0xFFFFFFFF;
+			int from_k = (int)from;
+			int to_k = (int)to;
 			int from_vertex = wavefrontToKey.get(from_k);
 			int to_vertex = wavefrontToKey.get(to_k);
 			pw.println("l "+from_vertex+" "+to_vertex);			

### 2008-05-06 tracing/NewAnalyzeTracings_.java
The basis of a new connectivity analysis class

@@ -1,0 +1,184 @@
+/* -*- mode: java; c-basic-offset: 8; indent-tabs-mode: t; tab-width: 8 -*- */
+
+package tracing;
+
+import java.awt.*;
+import java.util.Hashtable;
+import java.util.HashSet;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.io.*;
+import java.util.Enumeration;
+
+import amira.AmiraParameters;
+
+import ij.*;
+import ij.io.*;
+import ij.gui.*;
+import ij.plugin.PlugIn;
+
+import vib.transforms.OrderedTransformations;
+
+import util.FileAndChannel;
+import vib.oldregistration.Bookstein_FromMarkers;
+
+import util.BatchOpener;
+
+class ImagesFromLine {
+	
+	String lineName;
+	ArrayList<String> baseNames = new ArrayList<String>();
+	
+}
+
+public class NewAnalyzeTracings_ implements PlugIn {
+
+		
+	String [] labelNames = { "Ellipsoid Body",
+				 "Fan-Shaped Body",
+				 "Noduli",
+				 "Protocerebral Bridge" };
+		
+	int [] labelIndices =  { 9,
+				 11,
+				 10,
+				 12 };
+	
+      	public Connectivity buildGraph( String tracesObjFileName, String labelsFileName ) {
+		
+		Connectivity result = new Connectivity();		
+		int x, y, z;
+
+		/* First load the traces file:
+
+
+
+
+		 */
+
+
+		
+		/* This nonsense is all about loading the label files: */
+
+		ImagePlus labels = BatchOpener.openFirstChannel(labelsFileName);
+		if( labels == null )
+			throw new RuntimeException("Couldn't open labels file "+labelsFileName);
+		ImageStack labelStack=labels.getStack();
+
+		byte[][] label_data=new byte[templateDepth][];
+		for( z = 0; z < templateDepth; ++z )
+			label_data[z] = (byte[])labelStack.getPixels( z + 1 );
+		
+		AmiraParameters parameters = new AmiraParameters(labels);
+		int materials = parameters.getMaterialCount();
+		result.materialNames = new String[256];
+		result.materialNameToIndex = new Hashtable< String, Integer >();
+		for( int i = 0; i < materials; ++i ) {
+			result.materialNames[i] = parameters.getMaterialName(i);
[…]

### 2008-05-07 tracing/Load_Auto_Traces.java
Fix a loading bug and separated out the traces.obj loading code

@@ -31,25 +31,69 @@ import java.io.*;
 import java.awt.*;
 import java.util.StringTokenizer;
 
-public class Load_Auto_Traces implements PlugIn {
+public class Load_Auto_Traces implements PlugIn, TraceLoaderListener {
 
-	public int getNonNegativeIntAfterPrefix( String line, String prefix ) {
-		try {
-			String s=line.substring(prefix.length());
-			return Integer.parseInt(s);
-		} catch( NumberFormatException e ) {
-			return -1;
+	int width = -1, height = -1, depth = -1;
+	float spacing_x = Float.MIN_VALUE;
+	float spacing_y = Float.MIN_VALUE;
+	float spacing_z = Float.MIN_VALUE;
+
+	byte [][] values = null;
+	
+	public void gotVertex( int vertexIndex,
+			       float x_scaled, float y_scaled, float z_scaled,
+			       int x_image, int y_image, int z_image ) {
+
+		if( values == null ) {
+			if( width < 0 ||
+			    height < 0 ||
+			    depth < 0 ||
+			    spacing_x == Float.MIN_VALUE ||
+			    spacing_y == Float.MIN_VALUE ||
+			    spacing_z == Float.MIN_VALUE ) {
+
+				throw new RuntimeException("Some metadata was missing from the comments before the first vertex.");
+			}
+			values = new byte[depth][];
+			for( int z = 0; z < depth; ++z )
+				values[z] = new byte[width*height];
 		}
+
+		if( z_image >= depth ) {
+			System.out.println("z_image: "+z_image+" was too large for depth: "+depth);
+			System.out.println("z_scaled was: "+z_scaled);
+		}
+
+		values[z_image][y_image*width+x_image] = (byte)255;
 	}
 
-	public float getNonNegativeFloatAfterPrefix( String line, String prefix ) {
-		try {
-			String s=line.substring(prefix.length());
-			return Float.parseFloat(s);
-		} catch( NumberFormatException e ) {
-			return -1;
-		}
+	public void gotLine( int fromVertexIndex, int toVertexIndex ) {
+		// Do nothing...
 	}
+
+	public void gotWidth( int width ) {
+		this.width = width;
+	}
+
+	public void gotHeight( int height ) {
+		this.height = height;
+	}
+
+	public void gotDepth( int depth ) {
+		this.depth = depth;
+	}
+
[…]

### 2008-05-07 tracing/SinglePathsGraph.java

@@ -28,6 +28,16 @@ import java.util.Iterator;
 import java.io.*;
 import java.util.zip.GZIPOutputStream;
 
+import ij.*;
+import ij.plugin.*;
+import ij.gui.*;
+import ij.io.*;
+import ij.process.*;
+
+import java.io.*;
+import java.awt.*;
+import java.util.StringTokenizer;
+
 public class SinglePathsGraph {
 	
 	int width, height, depth;
@@ -128,6 +138,120 @@ public class SinglePathsGraph {
 		pw.close();
 	}
 
+	static public int getNonNegativeIntAfterPrefix( String line, String prefix ) {
+		try {
+			String s=line.substring(prefix.length());
+			return Integer.parseInt(s);
+		} catch( NumberFormatException e ) {
+			return -1;
+		}
+	}
+
+	static public float getNonNegativeFloatAfterPrefix( String line, String prefix ) {
+		try {
+			String s=line.substring(prefix.length());
+			return Float.parseFloat(s);
+		} catch( NumberFormatException e ) {
+			return -1;
+		}
+	}
+
+	static boolean loadWithListener( String inputFilename, TraceLoaderListener listener ) {
+
+		int width = -1, height = -1, depth = -1;
+		float spacing_x = Float.MIN_VALUE;
+		float spacing_y = Float.MIN_VALUE;
+		float spacing_z = Float.MIN_VALUE;
+
+		try {
+			String widthPrefix = "# width: ";
+			String heightPrefix = "# height: ";
+			String depthPrefix = "# depth: ";
+			String spacingXPrefix = "# spacing_x: ";
+			String spacingYPrefix = "# spacing_y: ";
+			String spacingZPrefix = "# spacing_z: ";
+
+			String vertexPrefix = "v ";
+			String linePrefix = "l ";
+			String groupPrefix = "g ";
+
+			int vertexIndex = 1;
+
+			BufferedReader br = new BufferedReader(new FileReader(inputFilename));
+			String lastLine;
+			while( null != (lastLine = br.readLine()) ) {
+				
+				if( lastLine.startsWith(widthPrefix) ) {
+					width = getNonNegativeIntAfterPrefix(lastLine, widthPrefix);
+					listener.gotWidth(width);
+				}
+				if( lastLine.startsWith(heightPrefix) ) {
+					height = getNonNegativeIntAfterPrefix(lastLine, heightPrefix);
[…]

### 2008-05-07 tracing/NewAnalyzeTracings_.java
Load the graph data, find all the edge points - no connection finding yet.

@@ -31,9 +31,8 @@ class ImagesFromLine {
 	
 }
 
-public class NewAnalyzeTracings_ implements PlugIn {
+public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 
-		
 	String [] labelNames = { "Ellipsoid Body",
 				 "Fan-Shaped Body",
 				 "Noduli",
@@ -44,59 +43,217 @@ public class NewAnalyzeTracings_ implements PlugIn {
 				 10,
 				 12 };
 	
-      	public Connectivity buildGraph( String tracesObjFileName, String labelsFileName ) {
+	class NewGraphNode {
+		public NewGraphNode() { }
+		public NewGraphNode(int x, int y, int z) {
+			this.x = x;
+			this.y = y;
+			this.z = z;
+		}
+		int x = -1;
+		int y = -1;
+		int z = -1;
+		NewGraphNode linkedTo [] = null;
+		public boolean equals(Object other) {
+			NewGraphNode o=(NewGraphNode)other;
+			return x == o.x && y == o.y && z == o.z;
+		}
+	}
+
+	public int positionToKey( int x, int y, int z ) {
+		return x + y * width + z * width * height;
+	}
+
+	int width = -1, height = -1, depth = -1;
+	float spacing_x = Float.MIN_VALUE;
+	float spacing_y = Float.MIN_VALUE;
+	float spacing_z = Float.MIN_VALUE;
+
+	ArrayList<NewGraphNode> verticesInObjOrder;
+	Hashtable<Integer,NewGraphNode> positionToNode = new Hashtable<Integer,NewGraphNode>();
+	int numberOfVertices = -1;
+	ArrayList<ArrayList<NewGraphNode>> links; 
+	
+	public void gotVertex( int vertexIndex,
+			       float x_scaled, float y_scaled, float z_scaled,
+			       int x_image, int y_image, int z_image ) {
 		
-		Connectivity result = new Connectivity();		
-		int x, y, z;
+		if( width < 0 ||
+		    height < 0 ||
+		    depth < 0 ||
+		    spacing_x == Float.MIN_VALUE ||
+		    spacing_y == Float.MIN_VALUE ||
+		    spacing_z == Float.MIN_VALUE ) {
 
-		/* First load the traces file:
+			throw new RuntimeException("Some metadata was missing from the comments before the first vertex.");
+		}
 
[…]

### 2008-05-07 tracing/NewAnalyzeTracings_.java
More progress

@@ -54,10 +54,31 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		int y = -1;
 		int z = -1;
 		NewGraphNode linkedTo [] = null;
+		@Override
 		public boolean equals(Object other) {
 			NewGraphNode o=(NewGraphNode)other;
 			return x == o.x && y == o.y && z == o.z;
 		}
+		@Override
+		public int hashCode( ) {
+			return x + y * (1 << 11) + z * (1 << 22);
+		}
+		@Override
+		public String toString( ) {
+			return "("+x+","+y+","+z+")";
+		}
+	}
+
+	Path findPath( NewGraphNode startPoint, NewGraphNode endPoint ) {
+		
+
+
+
+
+
+
+
+		return null;
 	}
 
 	public int positionToKey( int x, int y, int z ) {
@@ -212,10 +233,11 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		int [] greenValues = new int[materials];
 		int [] blueValues = new int[materials];
 
+		ArrayList<ArrayList<NewGraphNode>> allEdges = new ArrayList<ArrayList<NewGraphNode>>();
+
 		for( int i=0; i < materials; i++ ) {
-			
-			double[] c = parameters.getMaterialColor(i);
-			
+			allEdges.add(new ArrayList<NewGraphNode>());
+			double[] c = parameters.getMaterialColor(i);			
 			redValues[i] = (int)(255*c[0]);
 			greenValues[i] = (int)(255*c[1]);
 			blueValues[i] = (int)(255*c[2]);
@@ -227,7 +249,7 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 
 			System.out.println("   Dealing with label index "+labelIndex+", name: "+labelPrettyString);			
 
-			ArrayList<NewGraphNode> neuropilEdgePoints = new ArrayList<NewGraphNode>();
+			ArrayList<NewGraphNode> neuropilEdgePoints = allEdges.get(labelIndex);
 
 			for( int z = 0; z < depth; ++z )
 				for( int y = 0; y < height; ++y )
@@ -250,8 +272,57 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 							}
 						}
 					}
+
 			System.out.println("   Found "+neuropilEdgePoints.size()+" points on the edge of the "+labelPrettyString);
 		}
+
+		// Now start a search from each of these points trying
[…]

### 2008-05-07 tracing/NewAnalyzeTracings_.java
Crazy n-squared search from all end points to all endpoints

@@ -6,7 +6,9 @@ import java.awt.*;
 import java.util.Hashtable;
 import java.util.HashSet;
 import java.util.ArrayList;
+import java.util.Collection;
 import java.util.Iterator;
+import java.util.PriorityQueue;
 import java.io.*;
 import java.util.Enumeration;
 
@@ -43,7 +45,7 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 				 10,
 				 12 };
 	
-	class NewGraphNode {
+	class NewGraphNode implements Comparable {
 		public NewGraphNode() { }
 		public NewGraphNode(int x, int y, int z) {
 			this.x = x;
@@ -67,17 +69,225 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		public String toString( ) {
 			return "("+x+","+y+","+z+")";
 		}
+		// These members are only used for the search:
+		float g = Float.MIN_VALUE;
+		float h = Float.MIN_VALUE;
+		NewGraphNode previous = null;
+		public float distanceTo( NewGraphNode o ) {
+			float xdiff = x - o.x;
+			float ydiff = y - o.y;
+			float zdiff = z - o.z;
+			float distSq = xdiff*xdiff + ydiff*ydiff + zdiff*zdiff;
+			return (float) Math.sqrt( distSq );
+		}
+		public void setFrom( NewGraphNode o ) {
+			this.x = o.x;
+			this.y = o.y;
+			this.z = o.z;
+			this.linkedTo = o.linkedTo;
+		}	
+		double f() {
+			return g + h;
+		}	
+		@Override
+		public int compareTo( Object other ) {
+			NewGraphNode n = (NewGraphNode)other;
+			return Double.compare( f(), n.f() );
+		}
 	}
 
-	Path findPath( NewGraphNode startPoint, NewGraphNode endPoint ) {
+	ArrayList< NewGraphNode > makePath( NewGraphNode lastNode ) {       
 		
+		// System.out.println( "Trying to return result" );
+		
+		ArrayList< NewGraphNode > resultReversed = new ArrayList< NewGraphNode >();
+		NewGraphNode p = lastNode;
+		do {
+			resultReversed.add(p);
+			// System.out.println(  "adding "+p.toDotName());
+		} while( null != (p = p.previous) );
+		
+		ArrayList< NewGraphNode > realResult = new ArrayList< NewGraphNode >();
+		
[…]

### 2008-05-07 tracing/NewAnalyzeTracings_.java
Much faster just looking for desintation materials

@@ -124,12 +124,11 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		public ArrayList< NewGraphNode > path;      
 	}
 
-	PathWithLength findPath( NewGraphNode start, NewGraphNode end ) {
+	PathWithLength findPath( NewGraphNode start, int endMaterial ) {
 
 		// System.out.println("Starting path finding:");
 
 		int startMaterial = label_data[start.z][start.y*width+start.x];
-		int endMaterial = label_data[end.z][end.y*width+end.x];
 			
 		// First reset all the search parameters:
 		{
@@ -169,9 +168,11 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 			
 			// System.out.println( " Got node "+p.toDotName()+" from the queue" );
 			
-			// Has the route from the start found the goal?
-			
-			if( p.equals( end ) ) {
+			int pointMaterial = label_data[p.z][p.y*width+p.x];
+
+			// Has the route from the start found the goal?			
+
+			if( pointMaterial == endMaterial ) {
 				// System.out.println( "Found the goal! (from start to end)" );
 				ArrayList< NewGraphNode > path = makePath(p);
 				if( path == null )
@@ -190,7 +191,6 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 			
 			// Now look at all the neighbours...
 
-
 			// System.out.println("linkedTo "+p.linkedTo.length+" neigbours");
 			for( int i = 0; i < p.linkedTo.length; ++i ) {
 
@@ -198,25 +198,10 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 				float distance = p.distanceTo(neighbour);
 				int neighbourMaterial = label_data[neighbour.z][neighbour.y*width+neighbour.x];
 
-				// Ignore this neighbour if it's:
-				//   - if it's not the exterior material or end material
-				//   - of the original material
-				//   - of the end material and not the goal
+				// Ignore this neighbour if it's it's not of the exterior or end material
 
-				if( neighbourMaterial == 0 ) {
-					// Exterior, fine...
-					// System.out.println("   adding exterior point");
-				} else if( neighbourMaterial == startMaterial ) {
+				if( ! (neighbourMaterial == 0 || neighbourMaterial == endMaterial) )
 					continue;
[…]

### 2008-05-07 tracing/NewAnalyzeTracings_.java
Enable output of paths to a file via PathAndFillManager

@@ -121,7 +121,16 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		int endNeuropilRegion;
 
 		public double length;
-		public ArrayList< NewGraphNode > path;      
+		public ArrayList< NewGraphNode > path;
+
+		public Path toPath() {
+			Path p = new Path(path.size());
+			for( int i = 0; i < path.size(); ++i ) {
+				NewGraphNode n = path.get(i);
+				p.addPoint(n.x,n.y,n.z);
+			}
+			return p;
+		}
 	}
 
 	PathWithLength findPath( NewGraphNode start, int endMaterial ) {
@@ -355,7 +364,7 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 	byte[][] label_data;
 	String [] materialNames;
 
-      	public Connectivity buildGraph( File tracesObjFile, File labelsFile ) {
+      	public ArrayList<PathWithLength> buildGraph( File tracesObjFile, File labelsFile, File writePathsTo, File writeDotTo ) {
 
 		String tracesObjFileName = tracesObjFile.getAbsolutePath();
 		String labelsFileName = labelsFile.getAbsolutePath();
@@ -432,7 +441,7 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		int [] blueValues = new int[materials];
 
 		ArrayList<ArrayList<NewGraphNode>> allEdges = new ArrayList<ArrayList<NewGraphNode>>();
-
+		
 		for( int i=0; i < materials; i++ ) {
 			allEdges.add(new ArrayList<NewGraphNode>());
 			double[] c = parameters.getMaterialColor(i);			
@@ -440,6 +449,8 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 			greenValues[i] = (int)(255*c[1]);
 			blueValues[i] = (int)(255*c[2]);
 		}
+
+/* Find all the points on the edge of a neuropil regions: */
 
 		for( int a = 0; a < labelIndices.length; ++a ) {
 			int labelIndex = labelIndices[a];
@@ -474,6 +485,14 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 			System.out.println("   Found "+neuropilEdgePoints.size()+" points on the edge of the "+labelPrettyString);
 		}
 
+		// We'll store copies of these in a PathAndFillManager
+		// so that we can write out something that will be
[…]

### 2008-05-07 tracing/Path.java

@@ -48,7 +48,15 @@ public class Path implements Cloneable {
 	
 	public static final int PATH_START = 0;
 	public static final int PATH_END = 1;
+
+	// It's sometimes useful to give paths a name, but this will
+	// mostly be null:
+	String name;
 	
+	public void setName(String newName) {
+		this.name = newName;
+	}
+
 	public double getRealLength( double x_spacing, double y_spacing, double z_spacing ) {
 		double totalLength = 0;
 		for( int i = 1; i < points; ++i  ) {
@@ -728,7 +736,10 @@ public class Path implements Cloneable {
 	@Override
         public String toString() {
 		int n = size();
-		String result = "" + n + " points";
+		String result = "";
+		if( name != null )
+			result += "\"" + name + "\" ";
+		result += n + " points";
 		if( n > 0 ) {
 			result += " from " + x_positions[0] + ", " + y_positions[0] + ", " + z_positions[0];
 			result += " to " + x_positions[n-1] + ", " + y_positions[n-1] + ", " + z_positions[n-1];

### 2008-05-07 tracing/PathAndFillManager.java

@@ -89,6 +89,18 @@ public class PathAndFillManager extends DefaultHandler {
 		this.height = plugin.height;
 		this.depth = plugin.depth;
         }
+
+	public PathAndFillManager( int width, int height, int depth, float x_spacing, float y_spacing, float z_spacing, String spacing_units ) {
+		this();
+		this.x_spacing = x_spacing;
+		this.y_spacing = y_spacing;
+		this.z_spacing = z_spacing;
+		this.width = width;
+		this.height = height;
+		this.depth = depth;
+		if( spacing_units == null )
+			this.spacing_units = "unknown";
+	}
 	
 	int width;
 	int height;

### 2008-05-07 tracing/NewAnalyzeTracings_.java
Output a simple dot file for neato as well

@@ -364,6 +364,10 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 	byte[][] label_data;
 	String [] materialNames;
 
+	int [] redValues;
+	int [] greenValues;
+	int [] blueValues;
+
       	public ArrayList<PathWithLength> buildGraph( File tracesObjFile, File labelsFile, File writePathsTo, File writeDotTo ) {
 
 		String tracesObjFileName = tracesObjFile.getAbsolutePath();
@@ -429,16 +433,16 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		AmiraParameters parameters = new AmiraParameters(labels);
 		int materials = parameters.getMaterialCount();
 		materialNames = new String[256];
-		Hashtable<String,Integer> materialNameToIndex = new Hashtable< String, Integer >();
+		materialNameToIndex = new Hashtable< String, Integer >();
 		for( int i = 0; i < materials; ++i ) {
 			materialNames[i] = parameters.getMaterialName(i);
 			materialNameToIndex.put(materialNames[i],new Integer(i));
 			System.out.println("Material: "+i+" is "+materialNames[i]);
 		}
 		
-		int [] redValues = new int[materials];
-		int [] greenValues = new int[materials];
-		int [] blueValues = new int[materials];
+		redValues = new int[materials];
+		greenValues = new int[materials];
+		blueValues = new int[materials];
 
 		ArrayList<ArrayList<NewGraphNode>> allEdges = new ArrayList<ArrayList<NewGraphNode>>();
 		
@@ -524,6 +528,9 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 						continue;
 					}
 
+					route.startNeuropilRegion = labelIndex;
+					route.endNeuropilRegion = endM;
+
 					paths.add(route);
 					Path newPath=route.toPath();
 					newPath.setName(materialNames[labelIndex]+" to " +materialNames[endM]);
@@ -547,12 +554,49 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		}
 		
 		if( writeDotTo != null ) {
-			
-			
+			try {				
+				BufferedWriter out = new BufferedWriter(new FileWriter(writeDotTo.getAbsolutePath(),false));
+				
+				out.write( "graph G {\n" );
+				out.write( "        graph [overlap=scale,splines=true];\n");
[…]

### 2008-05-07 compare-stacks.rb
Old changes, possibly no longer relevant

@@ -5,7 +5,52 @@
 
 vib_directory=File.dirname(File.expand_path(__FILE__))
 
+require 'getoptlong'
+options = GetoptLong.new(
+  [ "--help", "-h", GetoptLong::NO_ARGUMENT ],
+  [ "--substring", "-s", GetoptLong::REQUIRED_ARGUMENT ]
+)
+
+program_name = "compare-stacks"
 memory="512m"
+
+def usage
+  print <<EOUSAGE
+Usage: #{program_name} [OPTION]...
+
+  -h, --help           Display this message and exit.
+  -s <SUBSTRING>, --substring=<SUBSTRING>
+                       Only use the images with <SUBSTRING> in their
+                       titles.
+  -k --keep-sources    Keep source images
+  -c --close-others    Close non-matching images
+EOUSAGE
+end
+
+substring = ""
+
+keep_sources = false
+close_others = false
+
+begin
+  options.each do |opt, arg|
+    case opt
+    when "--help"
+      usage
+      exit
+    when "--substring"
+      substring = arg
+    when "--keep-sources"
+      keep_sources = true
+    when "--close-others"
+      close_others = true
+    end
+  end
+rescue
+  print "Bad command line option: " + $! + "\n"
+  usage
+  exit
+end
 
 unless ARGV.length == 2
 	puts "Usage: compare-stacks <fileA> <fileB>"
@@ -28,11 +73,14 @@ end
 fileA=File.expand_path(fileA)
 fileB=File.expand_path(fileB)
 
+macro_options = "substring=#{substring}"
+macro_options += ",keep" if keep_sources
+macro_options += ",close" if close_others
+
 Dir.chdir( vib_directory ) {
 
-	result = system( "java", "-Xmx#{memory}", "-Dplugins.dir=.", "-jar", "../ImageJ/ij.jar", "-port0", fileA, fileB, "-eval", "run('Overlay Registered','');" )
+	result = system( "java", "-Xmx#{memory}", "-Dplugins.dir=.", "-jar", "../ImageJ/ij.jar", "-port0", fileA, fileB, "-eval", "run('Overlay Registered','#{macro_options}');" )
 	unless result
 		puts "Running ImageJ failed."
 	end
 }
-

### 2008-05-07 util/Overlay_Registered.java

@@ -8,28 +8,103 @@ import ij.gui.*;
 import ij.plugin.*;
 import ij.plugin.filter.*;
 
+import java.util.ArrayList;
+
 public class Overlay_Registered implements PlugIn {
+
+	public static float[] getValuesRange(ImagePlus imagePlus) {
 	
+		int stackSize      = imagePlus.getStackSize();
+		ImageStack stack      = imagePlus.getStack();
+		int bitDepth      = imagePlus.getBitDepth();
+
+		float minValue = Float.MAX_VALUE;
+		float maxValue = Float.MIN_VALUE;
+
+		int z;
+		for( z = 0; z < stackSize; ++z ) {
+			if( 8 == bitDepth) {
+				byte [] pixels = (byte[])stack.getPixels(z+1);
+				for( int i = 0; i < pixels.length; ++i ) {
+					int value = pixels[i] & 0xFF;
+					if( value > maxValue )
+						maxValue = value;
+					if( value < minValue )
+						minValue = value;
+				}
+			} else if( 16 == bitDepth ) {
+				short [] pixels = (short[])stack.getPixels(z+1);
+				for( int i = 0; i < pixels.length; ++i ) {
+					short value = pixels[i];
+					if( value > maxValue )
+						maxValue = value;
+					if( value < minValue )
+						minValue = value;
+				}
+			}
+		}
+
+		float [] result = new float[2];
+
+		result[0] = minValue;
+		result[1] = maxValue;
+
+		return result;
+	}
+
 	public void run(String ignored) {
-		
+	
+		String macroOptions = Macro.getOptions();
+
+		String mustHaveSubstring = "";
+
+		if (macroOptions != null) {
+			String value = Macro.getValue(macroOptions, "substring", null);
+			if( value != null ) {
+				mustHaveSubstring = value;
+			}
+			value = Macro.getValue(macroOptions, "keep", null);
+			if( value != null ) {
+				System.out.println("Got keep!: '"+value+"'");
+			}
+			value = Macro.getValue(macroOptions, "close", null);
+			if( value != null ) {
+				System.out.println("Got close!: '"+value+"'");
+			}
+		}
+
 		int[] wList = WindowManager.getIDList();
 		if (wList == null) {
 			IJ.error("No images are open.");
 			return;
 		}
-
-		String[] titles = new String[wList.length + 1];
-		for (int i = 0; i < wList.length; i++) {
[…]

### 2008-05-07 viewer_paper/paper.tex

--- 2008-04-16 viewer_paper/paper.tex
@@
 \begin{abstract}
-In this paper, we introduce a tool for visualising 3D data sets.
+In this paper, we introduce a tool for visualizing 3D datasets.
 Microscopy techniques in biology and medicine, like confocal microscopy, EM, MRI etc, often provide a stack of 2D images.
 For the investigation of these images, however, the two-dimensional view of the single slices is often not sufficient.
-We therefore developed a software for 3D visualisation, which allows to display the stacks optionally as surfaces, volume renderings or orthoslices.
-The software focuses on interactive, hardware-accelerated display of the stacks.
+We therefore developped a software for 3D visualisation, which allows to display the stacks optionally as surfaces, volume renderings or orthoslices.
+The software focusses on interactive, hardware-accelerated display of the stacks.
 Additionally, it offers possibilities for animation and movie recording, as well as an elaborate way to handle transformations of displayed bodies.
@@
 \item It is freely available and extensible.
-\item It was developed for simplicity of usage.
+\item It was developped for simplicity of usage.
 \item It is incorporated into ImageJ \cite{imagej}, a prominent image processing software for microscopy data.
@@
 \section{Introduction}
-Medical and biological imaging techniques often acquire 3D data sets in the form of stacks of 2D images.
+Medical and biological imaging techniques often acquire 3D datasets in the form of stacks of 2D images.
 Confocal microscopy for example can be used to cut a \textit{Drosophila} brain virtually into slices of a few $\mu$m thickness and scans it.
@@
 Its most relevant advantage however is its incorporation into ImageJ, which is one of the most wide-spread image processing packages in the field of microscopy.
-Users of ImageJ who already benefit from its numerous possibilities and plugins to process microscopic images, like filtering and deconvolution \cite{mbf}, can now visualise their data in 3D seamlessly, without the need of another software.
+Users of ImageJ who already benefit from its numerous possibilities and plugins to process microscopic images, like filtering and deconvolution \cite{mbf}, can now visualize their data in 3D seemlessly, without the need of another software.
 
-This paper is organised as follows: We first introduce briefly the technologies used for the implementation of the software.
+This paper is organized as follows: We first introduce briefly the technologies used for the implementation of the software.
 Subsequently, we show how the surface, volume and orthoslice views are created, as well as more information about the features of this software.
@@
 \textbf{ImageJ} \cite{imagej} is used as underlying platform.
-There are several reasons for choosing ImageJ for this task: Using an image processing platform in general has the advantage that the developer does not need to take care about low-level actions like opening/saving images etc, since functionality like that is given by the chosen platform.
+There are several reasons for choosing ImageJ for this task: Using an image processing platform in general has the advantage that the developper does not need to take care about low-level actions like opening/saving images etc, since functionality like that is given by the chosen platform.
 The choice fell on ImageJ since it is by far the most widely used image processing software in the biological and medical field \cite{mbf}.
@@
 \item It is freely available and open source.
-\item It consists of a compact core covering basic functionality, and is easily extensible to the user's needs by plugins.
+\item It consists of a compact core covering basic functionality, and is easily extendible to the user's needs by plugins.
 \item There exists a active and competent community, which both serves as support via online forums and as developers for plugins of which hundreds can be found on the internet.
@@
 \textbf{Java 3D.} To guarantee interactive 3D visualisation, hardware capabilities must be used optimally.
-Many 3D rendering tasks can nowadays be fulfilled by graphics cards, which frees the CPU and accelerates operations drastically.
+Many 3D rendering tasks can nowadays be fullfilled by graphics cards, which frees the CPU and accelerates operations drastically.
 Java 3D \cite{java3d} is a technology which allows to use these hardware capabilities from within Java.
@@
 \section{3D Visualisation}
-3D data sets can be represented in different ways. The options we implemented in this work are surfaces, volume renderings and orthoslices.
+3D datasets can be represented in different ways. The options we implemented in this work are surfaces, volume renderings and orthoslices.
 Starting point is a set of 2D slices which contain volumetric data, e.g.
@@
 A surface is generally understood as the border between a volumetric body and its environment.
-Given a volumetric 3D data set, constructing a corresponding surface is an ill-defined problem.
+Given a volumetric 3D dataset, constructing a corresponding surface is an ill-defined problem.
 In computer graphics, one often uses isosurfaces for this purpose.
@@
 To display volumetric data as a surface, the triangles defining the border between body and background have to be generated.
-One of the most commonly applied algorithms recruited for this purpose is the marching cubes algorithm \cite{lorensen87}, which we also utilised here.
+One of the most commonly applied algorithms recruted for this purpose is the marching cubes algoithm \cite{lorensen87}, which we also utilized here.
 
-A virtual cube "marches" through the volumetric stack and inserts triangles, according to the intensity values at the vertexes of the cube (see figure \ref{fig:surface}a).
+A virtual cube "marches" through the volumetric stack and inserts triangles, according to the intensity values at the vertices of the cube (see figure \ref{fig:surface}a).
 
-The resulting list of triangles are displayed, taking into account attributes like colour and transparency.
+The resulting list of triangles are displayed, taking into account attributes like color and transparency.
 For a smooth, realistic representation, vertex normals are calculated and used for shading.
@@
 \subsection{Volume rendering}
-Another method to visualise a set of 2D slices is volume rendering.
+Another method to visualize a set of 2D slices is volume rendering.
 The general idea is to put the 2D slices one behind another.
@@
 (A tutorial about volume rendering in Java 3D can be found at \cite{volrend}.)
-Figure \ref{fig:volume}c shows a volume rendering. To display coloured stacks, the transparency value of each voxel is calculated from the mean intensity of the red, green and blue colour component.
+Figure \ref{fig:volume}c shows a volume rendering. To display coloured stacks, the transparency value of each voxel is calculated from the mean intensity of the red, green and blue color component.
 
-To separate different channels in a double staining, our software allows to select the colour channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b.
+To separate different channels in a double staining, our software allows to select the color channels to be displayed (figures \ref{fig:volume}a and \ref{fig:volume}b.
 Additionally, it provides the functionality to edit volumes via virtual scissors (figure \ref{fig:volume}d).
@@
 \begin{itemize}
-\item colour: It is possible to either show a body in its original colour (as in the original set of the 2D slices) or to adjust its red, green and blue colour component arbitrarily.
+\item color: It is possible to either show a body in its original color (as in the original set of the 2D slices) or to adjust its red, green and blue color component arbitrarily.
 \item threshold: In case of a surface, this is the previously mentioned isovalue of the calculated isosurface.
@@
 \item transparency value
-\item channels: The colour channels of the original 2D slices, which should be used for displaying the 3D body (see figure \ref{fig:volume}a and \ref{fig:volume}b).
+\item channels: The color channels of the original 2D slices, which should be used for displaying the 3D body (see figure \ref{fig:volume}a and \ref{fig:volume}b).
 \end{itemize}
 
-The user can either set these properties when adding a body to the 3D universe, or she can change existing properties later on.
+The user can either set these properties when adding a body to the 3D universe, or she can change existing properties lateron.
 
@@
 For several reasons, it is desirable to show more than one body in the 3D world.
-Figure \ref{fig:multiple} shows as an example a 3D world with an isosurface of a fly brain whole-mount, together with the volume rendering of some coloured neuropiles (lobula, mushroom bodies and antennal lobes).
+Figure \ref{fig:multiple} shows as an example a 3D world with an isosurface of a fly brain wholemount, together with the volume rendering of some coloured neuropiles (lobula, mushroom bodies and antennal lobes).
 
@@
 
-This possibility to specify defined transformations on the level of individual bodies can for example be used to investigate and visualise the results of 3D image registration.
+This possibility to specify defined transformations on the level of individual bodies can for example be used to investigate and visualize the results of 3D image registration.
 Image registration is the transformation of a model image in such a way that it matches a reference image \cite{jenett2006}.
@@
 by ImageJ's avi exporting function.
-Video files like this one can be easily incorporated in PowerPoint presentations.
+Video files like this one can be easily incorporated in powerpoint presentations.
+
+\subsection{Landmark sets and registration}
+Each body in the 3D world may optionally own a landmark set, i.e.
+a set of named points.
+Creating landmark sets is very much facilitated in our viewer: Using ImageJ's point selection tool and clicking on a body initiates landmark creation.
+A list with all landmarks shows up, offering the possibility to rename or
+remove different points, or to highlight them in the 3D view. Landmark sets can
+be displayed or hidden, saved to disk and be reloaded again.
+
+Landmark sets are very helpful if one simply wants to label special positions in a tissue.
+However, they are also of particular use for landmark based registration.
+As pointed out above, image registration or alignment is very important for biomedical image analysis, e.g.
+for comparison or collocalisation studies.
+
+Many papers which aim for fully automated image registration.
+Depending on the image quality, however, fully automated methods offen don't obtained the desired results.
+On the other hand side, landmark creation in a 3D view, particularly using orthoslices, is so straightforward, that we implemented a registration algorithm based on manual landmark selection.
+At the moment, this is a rigid registration method first introduced in // TODO, but we aim to add more sophisticated techniqes in the future.
+
+After initiating registration via the menu, a dialog pops up, asking the user for a model image and a reference image (the one the model is aligned to).
+Subsequently it guides the user through the process of landmark creation of both reference and model image, and calculates a global rigid transformation which best transforms the landmark set of the model image on that of the reference image.
+The transformed model can optionally be exported into a stack of 2D slices and stored as an ordinary tiff image.
 
@@
 As mentioned previously, ImageJ consists of a tight core covering basic image processing functionality.
-This core can be extended individually by installing additional plugins, either developed on one's own or obtained from third parties.
+This core can be extended individually by installing additional plugins, either developped on one's own or obtained from third parties.
 To meet ImageJ's structure, the 3D viewer was also implemented as a plugin. It can easily be installed by downloading it and placing it into ImageJ's plugins directory.
@@
 In case of an isosurface, the geometry holds the triangle mesh, in the case of a volume rendering or orthoslices, it holds a set of rectangles.
-\item Appearance: Defines several attributes of a Shape3D which influence its display, like colour, transparency, textures etc.
+\item Appearance: Defines several attributes of a Shape3D which influence its display, like color, transparency, textures etc.
 \end{itemize}
@@
 \item[TG translate:] TransformGroup for global translation. Acts on user interaction when no body is selected.
-\item[TG center:] TransformGroup for centring displayed bodies in the universe. Each time updated when a new body is added to the universe.
+\item[TG center:] TransformGroup for centering displayed bodies in the universe. Each time updated when a new body is added to the universe.
 \item[BG scene:] BranchGroup which holds a subtree for each displayed body.
@@
 \bibliographystyle{plain}
-\bibliography{bibliography}
+\bibliography{../bibliography}
 
@@
 A cube is moved to the through the volume.
-The vertexes of the cube are given by 8 neighbouring voxel values.
+The vertices of the cube are given by 8 neighbouring voxel values.
 Dependent on these 8 values, triangles are inserted at each position (a).
@@
 Cy3, a red dye, was used to stain some specific brain regions.
-Where both dyes coincide, the tissue shows up yellow.
+Where boths dyes coincide, the tissue shows up yellow.
 The viewer allows to specify the channels which should be displayed.
@@
 \caption{Multiple bodies.
-The isosurface shown in figure \ref{fig:surface} is shown here in light grey and a transparency value of 0.5.
+The isosurface shown in figure \ref{fig:surface} is shown here in lightgray and a transparency value of 0.5.
 Additionally, the volume rendering of another stack is displayed, which shows lobula (orange), mushroom bodies (magenta) and antennal lobes (blue) of the same brain.}

### 2008-05-08 vib/Fit_Sphere.java
Fixed whitespaces problems

@@ -37,12 +37,12 @@ import vib.app.FileGroup;
  * @author mark
  */
 public class Fit_Sphere implements PlugIn {
-	
+
 	public void run(String ignored) {
-		
+
 		String averagedNC82Path = "/Users/mark/central-complex-complete-vib-protocol/output_1/71yAAeastmost.tif";
 		String normalizedRootDirectory = "/Users/mark/central-complex-complete-vib-protocol/normalized/";
-		
+
 		int[][][] pointsInShells =
 {{
 { 226, 450, 50 },
@@ -70,7 +70,7 @@ public class Fit_Sphere implements PlugIn {
 }};
 
 
-/*		
+/*
 		int[][][] pointsInShells = { { {249, 365, 43},
 					       {216, 454, 43},
 					       {244, 552, 43},
@@ -87,57 +87,57 @@ public class Fit_Sphere implements PlugIn {
 					       {264, 450, 56},
 					       {297, 566, 56} }
 		};
-*/		
-		
+*/
+
 		/*
 		   These next two points are those that define the angle
 		   swepping through the fan-shaped body that we consider */
-		
+
 		int [] pointInferiorLowZ = { 294 ,466, 43 };
 		int [] pointInferiorHighZ = { 294 ,466, 49 };
-		
+
 		int [] pointSuperiorLowZ = { 234, 484, 42 };
 		int [] pointSuperiorHighZ = { 234, 484, 50 };
 
-		/* We ignore the y co-ordinates of these and 
+		/* We ignore the y co-ordinates of these and
 		   exclude those outside the angle in the XZ
 		   plane. */
-				    
+
 		ImagePlus[] channels = BatchOpener.open(averagedNC82Path);
 		if (channels == null) {
 			IJ.error("Couldn't open: " + averagedNC82Path);
 		}
-		
+
 		ImagePlus imp = channels[0];
-		
+
 		Calibration c = channels[0].getCalibration();
-		
+
 		double x_spacing = c.pixelWidth;
 		double y_spacing = c.pixelHeight;
 		double z_spacing = c.pixelDepth;
-		
+
 		int width = imp.getWidth();
 		int height = imp.getHeight();
 		int depth = imp.getStackSize();
 
 		// Scale the limit points with the calibration...
-		
+
 		double [] pointMaxZDefiningAngle = new double[3];
 		double [] pointMinZDefiningAngle = new double[3];
-				
+
[…]

### 2008-05-09 run.sh
Added another special case for another computer I use...

@@ -31,6 +31,9 @@ rioja|cognac.brookfield.local)
 claret)
 	MEM=350m
 	;;
+wiggle)
+	MEM=950m
+	;;
 esac
 
 # Profiling stuff (see http://jiprof.sourceforge.net/)

### 2008-05-09 tracing/NewAnalyzeTracings_.java
Use landmark point based registration to find unlabelled neuropil regions

@@ -18,6 +18,7 @@ import ij.*;
 import ij.io.*;
 import ij.gui.*;
 import ij.plugin.PlugIn;
+import ij.process.ByteProcessor;
 
 import vib.transforms.OrderedTransformations;
 
@@ -34,16 +35,15 @@ class ImagesFromLine {
 }
 
 public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
-
-	String [] labelNames = { "Ellipsoid Body",
-				 "Fan-Shaped Body",
-				 "Noduli",
-				 "Protocerebral Bridge" };
 		
-	int [] labelIndices =  { 9,
-				 11,
-				 10,
-				 12 };
+	int [] labelIndices =  { 7,     // mushroom_body_r
+				 8,     // mushroom_body_l
+				 9,     // ellipsoid_body
+				 11,    // noduli
+				 10,    // fan_shaped_body
+				 12,    // protocerebral_bridge
+				 15,    // antennal_lobe_r
+				 16 };  // antennal_lobe_l
 	
 	class NewGraphNode implements Comparable {
 		public NewGraphNode() { }
@@ -74,9 +74,9 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		float h = Float.MIN_VALUE;
 		NewGraphNode previous = null;
 		public float distanceTo( NewGraphNode o ) {
-			float xdiff = x - o.x;
-			float ydiff = y - o.y;
-			float zdiff = z - o.z;
+			float xdiff = (x - o.x) * spacing_x;
+			float ydiff = (y - o.y) * spacing_y;
+			float zdiff = (z - o.z) * spacing_y;
 			float distSq = xdiff*xdiff + ydiff*ydiff + zdiff*zdiff;
 			return (float) Math.sqrt( distSq );
 		}
@@ -88,7 +88,7 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 		}	
 		double f() {
 			return g + h;
-		}	
+		}
 		@Override
 		public int compareTo( Object other ) {
 			NewGraphNode n = (NewGraphNode)other;
@@ -205,6 +205,8 @@ public class NewAnalyzeTracings_ implements PlugIn, TraceLoaderListener {
 
 				NewGraphNode neighbour = p.linkedTo[i];
 				float distance = p.distanceTo(neighbour);
+				if( neighbour.z == 118 || (neighbour.y*width + neighbour.x) == 118 )
[…]

### 2008-05-10 util/Overlay_Registered.java
Fix a compilation error

@@ -62,7 +62,6 @@ public class Overlay_Registered implements PlugIn {
                 String titleSubstring = null;
                 boolean closeAllOthers = false;
                 
-                String macroOptions = Macro.getOptions();
 		if (macroOptions != null) {
 			String value = Macro.getValue(macroOptions, "substring", null);
 			if( value != null ) {
@@ -91,7 +90,6 @@ public class Overlay_Registered implements PlugIn {
 		ArrayList<String> matchingTitles = new ArrayList<String>();
 		ArrayList<ImagePlus> matchingImages = new ArrayList<ImagePlus>();
 
-                String [] matchingTitles=new String[wList.length];
                 ImagePlus [] matchingImagePlus=new ImagePlus[wList.length];
                 ImagePlus [] allImages=new ImagePlus[wList.length];
                 

### 2008-05-10 features/ComputeCurvatures.java
Fixed whitespace and a loop counter bug

@@ -26,30 +26,30 @@
  *
  * @author Stephan Preibisch
  * @version 1.0
- * 
+ *
  * Change in this version (Mark Longair):
- * 
+ *
  * - Made the top level plugin a wrapper for this class so that
  *   "features" package so that it can be used by classes in other
  *   packages.
- * 
+ *
  * - Now implements Runnable, with the void run() method creating
  *   the Gaussian and reporting progress via an optional callback.
  *   (If used in this way you need to use the constructor where you
  *   supply an ImagePlus, sigma and an optional callback.
- * 
+ *
  * - Switched to using Johannes's JacobiDouble class instead of the
  *   Jama classes, so we don't introduce an additional dependency.
  *   It's about 15% faster with JacobiDouble, and presumbly that could
  *   be faster again wtih JacobiFloat.
- * 
+ *
  * - Added ordering of the eigenvalues (optionally on absolute
  *   values).
- * 
+ *
  * TODO:
- * 
+ *
  *   Use calibration information.
- * 
+ *
  */
 
 package features;
@@ -300,17 +300,17 @@ public class ComputeCurvatures implements Runnable
     /* Generate the Gaussian based on this.imp and this.sigma */
 
     public void setup( ) {
-        
+
         try {
-            
+
             if( imp == null ) {
                 IJ.error("BUG: imp should not be null - are you using the right constructor?");
                 return;
             }
-            
+
             if( callback != null )
                 callback.proportionDone( 0.0 );
-            
+
             if (imp.getStackSize() > 1)
                 {
                     // IJ.log("3D");
@@ -328,10 +328,10 @@ public class ComputeCurvatures implements Runnable
                     if (data == null)
                         return;
                 }
-            
+
             boolean computeGauss = true;
             boolean showGauss = false;
-            
+
             // IJ.log("Computing Gauss image");
             if (_3D)
                 {
@@ -355,24 +355,24 @@ public class ComputeCurvatures implements Runnable
                     if (showGauss)
                         FloatArrayToImagePlus((FloatArray2D)data, "Gauss image", 0, 255).show();
                 }
-            
+
             if( callback != null )
                 callback.proportionDone( 1.0 );
 
-            
+
         } catch( OutOfMemoryError e ) {
-        
+
             long requiredMiB = ( imp.getWidth() *
                                  imp.getHeight() *
                                  imp.getStackSize() * 4 ) / (1024 * 1024);
[…]

### 2008-05-10 features/ComputeCurvatures.java
The tracer now works on single slices as well as stacks

@@ -378,15 +378,62 @@ public class ComputeCurvatures implements Runnable
 
     }
 
-    public void hessianEigenvaluesAtPoint( int x,
-                                           int y,
-                                           int z, // zero-indexed
-                                           boolean orderOnAbsoluteSize,
-                                           double [] result, /* should be 3 elements */
-                                           boolean fixUp ) {
-                                                                                                                 
+
+    public void hessianEigenvaluesAtPoint2D( int x,
+                                             int y,
+                                             boolean orderOnAbsoluteSize,
+                                             double [] result, /* should be 2 elements */
+                                             boolean fixUp ) {
+
+        if( _3D ) {
+            IJ.error("hessianEigenvaluesAtPoint2D( x, y, z, ... ) is only for 2D data.");
+            return;
+        }
+
+        FloatArray2D data2D = (FloatArray2D)data;
+
+        if( fixUp ) {
+
+            if( x == 0 )
+                x = 1;
+            if( x == (data2D.width - 1) )
+                x = data2D.width - 2;
+
+            if( y == 0 )
+                y = 1;
+            if( y == (data2D.height - 1) )
+                y = data2D.height - 2;
+        }
+
+        double [][] hessianMatrix = computeHessianMatrix2D(data2D, x, y, sigma);
+        double [] eigenValues = computeEigenValues(hessianMatrix);
+
+        // Don't assume that these are ordered.
+
+        double e0 = eigenValues[0];
+        double e1 = eigenValues[1];
+
+        double e0c = orderOnAbsoluteSize ? Math.abs( e0 ) : e0;
+        double e1c = orderOnAbsoluteSize ? Math.abs( e1 ) : e1;
+
+        if( e0c <= e1c ) {
+            result[0] = e0;
+            result[1] = e1;
+        } else {
+            result[0] = e1;
+            result[1] = e0;
+        }
+    }
+
+    public void hessianEigenvaluesAtPoint3D( int x,
+                                             int y,
+                                             int z, // zero-indexed
+                                             boolean orderOnAbsoluteSize,
+                                             double [] result, /* should be 3 elements */
+                                             boolean fixUp ) {
+
         if( ! _3D ) {
-            IJ.error("hessianEigenvaluesAtPoint( x, y, z, ... ) is only for 3D data.");
+            IJ.error("hessianEigenvaluesAtPoint3D( x, y, z, ... ) is only for 3D data.");
             return;
         }
 

### 2008-05-10 features/Surfaceness_.java

@@ -43,10 +43,10 @@ public class Surfaceness_ implements PlugIn, GaussianGenerationCallback {
                 for( int y = 1; y < height - 1; ++y ) {
                     for( int x = 1; x < width - 1; ++x ) {
 
-                        c.hessianEigenvaluesAtPoint( x, y, z,
-                                                     true, // order absolute
-                                                     evalues,
-                                                     false );
+                        c.hessianEigenvaluesAtPoint3D( x, y, z,
+                                                       true, // order absolute
+                                                       evalues,
+                                                       false );
 
                         int index = y * width + x;
 

### 2008-05-10 features/Tubeness_.java

@@ -38,10 +38,10 @@ public class Tubeness_ implements PlugIn, GaussianGenerationCallback {
 				for (int y = 1; y < height - 1; ++y) {
 					for (int x = 1; x < width - 1; ++x) {
 
-						c.hessianEigenvaluesAtPoint(x, y, z,
-						    true, // order absolute
-						    evalues,
-						    false);
+						c.hessianEigenvaluesAtPoint3D(x, y, z,
+                                                      true, // order absolute
+                                                      evalues,
+                                                      false);
 
 						int index = y * width + x;
 

### 2008-05-10 tracing/Albert_Test.java

@@ -85,7 +85,7 @@ public class Albert_Test implements PlugIn {
                 // interval is up, so don't set it too high.
 		long reportEveryMilliseconds = 3000;
 		
-		TracerThread tracer = new TracerThread(imagePlus, timeoutSeconds, reportEveryMilliseconds, start_x, start_y, start_z, goal_x, goal_y, goal_z, reciprocal, hessian, null, hessian != null );
+		TracerThread tracer = new TracerThread(imagePlus, timeoutSeconds, reportEveryMilliseconds, start_x, start_y, start_z, goal_x, goal_y, goal_z, reciprocal, depth == 1, hessian, null, hessian != null );
 		
 		System.out.println("Running tracer...");
 		tracer.run();

### 2008-05-10 tracing/Simple_Neurite_Tracer.java

@@ -673,6 +673,7 @@ public class Simple_Neurite_Tracer extends ThreePanes
 			y_end,
 			z_end,
 			true, // reciprocal
+			singleSlice,
 			(hessianEnabled ? hessian : null),
 			tubeness,
 			hessianEnabled );
@@ -908,6 +909,8 @@ public class Simple_Neurite_Tracer extends ThreePanes
 	
         TextWindow helpTextWindow;
 	
+	boolean singleSlice;
+
         // HessianAnalyzer hessianAnalyzer;
         ArchiveClient archiveClient;
 	
@@ -936,13 +939,9 @@ public class Simple_Neurite_Tracer extends ThreePanes
                                 return;
                         }
 
-			if( (currentImage.getWidth() < 2) ||
-                            (currentImage.getHeight() < 2) ||
-                            (currentImage.getStackSize() < 2) ) {
-                                IJ.error( "There must be at least two sample points in each dimension" );
-                                return;
-                        }
-			
+			if( currentImage.getStackSize() == 1 )
+				singleSlice = true;
+
                         if( currentImage.getType() != ImagePlus.GRAY8 ) {
                                 IJ.error("This plugin only works on 8 bit images at the moment.");
                                 return;
@@ -972,39 +971,44 @@ public class Simple_Neurite_Tracer extends ThreePanes
 
 			{
 				String originalFileName=file_info.fileName;
-				int lastDot=originalFileName.lastIndexOf(".");
-				String beforeExtension=originalFileName.substring(0, lastDot);
-				String tubesFileName=beforeExtension+".tubes.tif";
-				ImagePlus tubenessImage = null;
-				File tubesFile=new File(file_info.directory,tubesFileName);
-				System.out.println("Testing for the existence of "+tubesFile.getAbsolutePath());
-				if( tubesFile.exists() ) {
-					IJ.showStatus("Loading tubes file.");
-					tubenessImage=BatchOpener.openFirstChannel(tubesFile.getAbsolutePath());
-					System.out.println("Loaded the tubeness file");
-					if( tubenessImage == null ) {
-						IJ.error("Failed to load tubes image from "+tubesFile.getAbsolutePath()+" although it existed");
-						return;
-					}
-					if( tubenessImage.getType() != ImagePlus.GRAY32 ) {
[…]

### 2008-05-10 tracing/TracerThread.java

@@ -4,19 +4,19 @@
 
 /*
   This file is part of the ImageJ plugin "Simple Neurite Tracer".
-  
+
   The ImageJ plugin "Simple Neurite Tracer" is free software; you
   can redistribute it and/or modify it under the terms of the GNU
   General Public License as published by the Free Software
   Foundation; either version 3 of the License, or (at your option)
   any later version.
-  
+
   The ImageJ plugin "Simple Neurite Tracer" is distributed in the
   hope that it will be useful, but WITHOUT ANY WARRANTY; without
   even the implied warranty of MERCHANTABILITY or FITNESS FOR A
   PARTICULAR PURPOSE.  See the GNU General Public License for more
   details.
-  
+
   You should have received a copy of the GNU General Public License
   along with this program.  If not, see <http://www.gnu.org/licenses/>.
 */
@@ -33,57 +33,59 @@ import ij.ImagePlus;
  * doing an A* search with a choice of distance measures. */
 
 public class TracerThread extends SearchThread {
-	
+
         int start_x;
         int start_y;
         int start_z;
         int goal_x;
         int goal_y;
         int goal_z;
-	
+
         boolean reciprocal;
-	
+
         ComputeCurvatures hessian;
-	
+
         Path result;
-	
-	
+
+
 	@Override
 	protected boolean atStart( int x, int y, int z ) {
 		return (x == start_x) && (y == start_y) && (z == start_z);
 	}
-	
+
 	@Override
 	protected boolean atGoal( int x, int y, int z ) {
 		return (x == goal_x) && (y == goal_y) && (z == goal_z);
 	}
-	
+
 	@Override
 	protected double minimumCostPerUnitDistance( ) {
-		
+
 		double minimum_cost;
-		
+
                 if( hessian == null ) {
-			
+
                         minimum_cost = reciprocal ? ( 1 / 255.0 ) : 1;
-			
+
                 } else {
-			
+
                         // result = 1E-4;  /* for the ratio of e0/e1 */
                         // result = 0.002; // 1;  /* for e1 - e0 */
-			
+
                         minimum_cost = 1 / 60.0;
                 }
-		
+
 		return minimum_cost;
 	}
 
-	
+
 	float [][] tubeness;
 	boolean useHessian;
 
+	boolean singleSlice;
+
         /* If you specify 0 for timeoutSeconds then there is no timeout. */
-	
+
         public TracerThread( ImagePlus imagePlus,
 			     int timeoutSeconds,
[…]

### 2008-05-10 features/Surfaceness_.java
Make it an error to look for Surfaceness of a 2D image

@@ -19,6 +19,11 @@ public class Surfaceness_ implements PlugIn, GaussianGenerationCallback {
 
         original = WindowManager.getCurrentImage();
 
+        if( original.getStackSize() == 1 ) {
+            IJ.error("It only makes sense to look for Sufaceness of 3D images (stacks)>");
+            return;
+        }
+
         ComputeCurvatures c = new ComputeCurvatures( original, 1.0, this  );
 
         c.run();

### 2008-05-10 features/Tubeness_.java
Change Tubeness_ so that it works on 2D images (single slices) Some whitespace fixes to that file as well

@@ -1,5 +1,5 @@
-/*
- */
+/* -*- mode: java; c-basic-offset: 8; indent-tabs-mode: t; tab-width: 8 -*- */
+
 package features;
 
 import ij.*;
@@ -12,7 +12,7 @@ import ij.process.*;
 public class Tubeness_ implements PlugIn, GaussianGenerationCallback {
 
 	public ImagePlus generateTubenessImage(ImagePlus original) {
-	
+
 		Calibration calibration=original.getCalibration();
 
 		ComputeCurvatures c = new ComputeCurvatures(original, 1.0, this);
@@ -28,46 +28,85 @@ public class Tubeness_ implements PlugIn, GaussianGenerationCallback {
 
 		double[] evalues = new double[3];
 
-		for (int z = 0; z < depth; ++z) {
-
-			System.out.println("Working on slice: " + z);
+		if( depth == 1 ) {
 
 			float[] slice = new float[width * height];
 
-			if ((z >= 1) && (z < depth - 1)) {
-				for (int y = 1; y < height - 1; ++y) {
-					for (int x = 1; x < width - 1; ++x) {
+			for (int y = 1; y < height - 1; ++y) {
+				for (int x = 1; x < width - 1; ++x) {
 
-						c.hessianEigenvaluesAtPoint3D(x, y, z,
-                                                      true, // order absolute
-                                                      evalues,
-                                                      false);
+					c.hessianEigenvaluesAtPoint2D(x, y,
+								      true, // order absolute
+								      evalues,
+								      false);
 
-						int index = y * width + x;
+					int index = y * width + x;
 
-						if ((evalues[1] >= 0) || (evalues[2] >= 0)) {
+					if ((evalues[1] >= 0) || (evalues[0] >= 0)) {
 
-							// If either of the two principle eigenvalues
-							// is positive then the curvature is in the
-							// wrong direction - towards higher
-							// instensities rather than lower.
+						// If either of the two principle eigenvalues
+						// is positive then the curvature is in the
+						// wrong direction - towards higher
+						// instensities rather than lower.
 
-							slice[index] = 0;
+						slice[index] = 0;
 
-						} else {
+					} else {
 
-							slice[index] = (float) Math.sqrt(evalues[2] * evalues[1]);
+						slice[index] = (float) Math.abs(evalues[1]);
 
-						}
 					}
 				}
+				IJ.showProgress(1 / (double) height);
 			}
[…]
